---
title: "New full MD models of Textbook English"
author: "Elen Le Foll"
date: "09/11/2021"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
  editor_options: 
    markdown: 
      wrap: sentence
bibliography: "packages.bib"
nocite: '@*'
      
---

This script is part of the Online Appendix to my PhD thesis.

Please cite as: 
Le Foll, Elen. 2022. Textbook English: A Corpus-Based Analysis of the Language of EFL textbooks used in Secondary Schools in France, Germany and Spain. PhD thesis. Osnabrück University.

For more information, see: https://elenlefoll.github.io/TextbookEnglish/

Please note that the plot dimensions in this notebook have been optimised for the print version of the thesis.


# Set-up

*Built with R `r getRversion()`*  

```{r setup, include=TRUE, results = "hide", warning = FALSE, message = FALSE}

knitr::opts_chunk$set(echo = TRUE, paged.print=TRUE, fig.width = 10, warning=FALSE)

library(car) # For data wrangling
library(caret) # For its confusion matrix function
#library(clipr) # For quick exports to other programme
library(cowplot)
library(DescTools) # For 95% CI
library(dplyr)
library(emmeans)
#library(FactoMineR) # For Shiny app
library(factoextra) # For circular graphs of variables
library(forcats) # For the fct_relevel function
library(here) # For dynamic file paths
library(ggplot2) 
library(ggthemes) # For theme of factoextra plots
library(lme4) # For linear regression modelling
library(patchwork) # To create figures with more than one plot
#library(pca3d) # For 3-D plots (not rendered in html knit)
library(PCAtools) # For nice biplots of PCA results
library(purrr) # For data wrangling
library(psych) # For various useful stats function
library(sjPlot) # For model plots and tables
library(suffrager) # For pretty feminist colour palettes :)
library(visreg) # For plots of interaction effects

source(here("R_rainclouds.R")) # For geom_flat_violin rainplots

```


## PCA and 3-D plots

Import full data (see 7_Ref_data_prep.Rmd for data preparation steps)

```{r import-data}

data <- readRDS(here("FullMDA", "dataforPCA.rds")) 

data %>% 
  filter(Series=="NGL") %>% 
  group_by(Series, Level) %>% 
  summarise(wordcount = sum(Words))

```

This chunk can be used to perform the MDA on various subsets of the data.

```{r import-subset-data-for-PCA, eval=FALSE}

# Full dataset
data <- readRDS(here("FullMDA", "dataforPCA.rds")) 

# Subset of the data that excludes levels A and B textbooks
data <- readRDS(here("FullMDA", "dataforPCA.rds")) %>%
  filter(Level !="A" & Level != "B") %>%
  droplevels()
summary(data$Level)

# Subset of the data to only include one Country subcorpus of the TEC
data <- readRDS(here("FullMDA", "dataforPCA.rds")) %>%
  #filter(Country != "France" & Country != "Germany") %>% # Spain only
  #filter(Country != "France" & Country != "Spain") %>% # Germany only
  filter(Country != "Spain" & Country != "Germany") %>% # France only
  droplevels()
summary(data$Country)

# Perform PCA on random subset of the data to test the stability of the solution. Re-running this line will generate a new subset of 2/3 of the texts randomly sampled.
set.seed(13) # Seed used for the biplot printed in Chapter 8
data <- readRDS(here("FullMDA", "dataforPCA.rds")) %>%
  slice_sample(n = 4980*0.6, replace = FALSE)
nrow(data)
data$Filename[1:4]

```

This chunk is used to create the 3-D plots. These cannot be rendered in the html knit.

```{r PCA-3D-plots, eval=FALSE}

colnames(data)
pca <- prcomp(data[,9:ncol(data)], scale.=FALSE) # All quantitative variables
register <- factor(data[,"Register"]) 
corpus <- factor(data[,"Corpus"])
subcorpus <- factor(data[,"Subcorpus"])

summary(register)
summary(corpus)
summary(subcorpus)
summary(pca)

# 3-D plot

colours <- suf_palette(name = "london", n = 6, type = "continuous") 
colours2 <- suf_palette(name = "classic", n = 5, type = "continuous") 
colours <- c(colours, colours2[c(2:4)]) # Nine colours range
col6 <- colours[c(6,5,4,7,9,2)] # Good order for PCA
scales::show_col(col6)

col6 <- c("#F9B921", "#A18A33", "#722672", "#BD241E", "#267226", "#15274D")
names(col6) <- c("Textbook Conversation", "Textbook Fiction", "Textbook Informative", "Spoken BNC2014 Ref.", "Youth Fiction Ref.", "Info Teens Ref.")
shapes6 <- c(rep("cube", 3),rep("sphere", 3))
names(shapes6) <- c("Textbook Conversation", "Textbook Fiction", "Textbook Informative", "Spoken BNC2014 Ref.", "Youth Fiction Ref.", "Info Teens Ref.")

pca3d(pca, group = subcorpus, 
      components = 1:3,
      #components = 4:6,
      show.plane=FALSE,
      col = col6,
      shape = shapes6,
      radius = 0.7,
      legend = "right")

## Looking at all three Textbook English registers in one colour

col4 <- colours[c(1,3,7,9)]
col4 <- c("#EA7E1E", "#15274D", "#BD241E", "#267226")

names(col4) <- c("Textbook.English", "Informative.Teens", "Spoken.BNC2014", "Youth.Fiction")
shapes4 <- c("cube", rep("sphere", 3))
names(shapes4) <- c("Textbook.English", "Informative.Teens", "Spoken.BNC2014", "Youth.Fiction")

pca3d(pca, group = corpus, 
      show.plane=FALSE,
      components = 1:3,
      col = col4,
      shape = shapes4,
      radius = 0.7,
      legend = "right")

```

## PCA biplots

```{r PCAtools-pairsplot, fig.height=30, fig.width=25, warning=FALSE}

data2 <- data %>% 
  mutate(Source = recode_factor(Corpus, Textbook.English = "Textbook English (TEC)", Informative.Teens = "Reference corpora", Spoken.BNC2014 = "Reference corpora", Youth.Fiction = "Reference corpora")) %>% 
  mutate(Corpus = fct_relevel(Subcorpus, "Info Teens Ref.", after = 9)) %>%
  relocate(Source, .after = "Corpus") %>% 
  droplevels(.)

colnames(data2)
data2meta <- data2[,1:9]
rownames(data2meta) <- data2meta$Filename
data2meta <- data2meta %>% select(-Filename)
head(data2meta)

rownames(data2) <- data2$Filename
data2num <- as.data.frame(base::t(data2[,10:ncol(data2)]))
data2num[1:5,1:5] # Check data frame format is correct

p <- PCAtools::pca(data2num, 
         metadata = data2meta,
         scale = FALSE)

p$variance[1:6]
sum(p$variance[1:4])

# For five TEC registers
# colkey = c(`Spoken BNC2014 Ref.`="#BD241E", `Info Teens Ref.`="#15274D", `Youth Fiction Ref.`="#267226", `Textbook Fiction`="#A18A33", `Textbook Conversation`="#F9B921", `Textbook Informative` = "#722672", `Textbook Instructional` = "grey", `Textbook Personal` = "black")

# For three TEC registers
summary(data2$Corpus)
colkey = c(`Spoken BNC2014 Ref.`="#BD241E", `Info Teens Ref.`="#15274D", `Youth Fiction Ref.`="#267226", `Textbook Fiction`="#A18A33", `Textbook Conversation`="#F9B921", `Textbook Informative` = "#722672")

#summary(data2$Source)
#shapekey = c(`Textbook English (TEC)`=6, `Reference corpora`=1)

summary(data2$Level)
shapekey = c(A=1, B=2, C=6, D=0, E=5, `Ref.`=4)

## Very slow, open in zoomed out window!
# Add legend manually? Yes (take it from the biplot code below), let's not waste too much time, here. Or use Evert's mvar.pairs plot (though that annoyingly requires manual axis annotation so not really better!)

#png(here("plots", "PCA_3Ref_pairsplot.png"), width = 25, height= 40, units = "cm", res = 300)
PCAtools::pairsplot(p,
                 triangle = FALSE,
                 components = 1:4,
                 ncol = 2,
                 nrow = 3,
                 pointSize = 1,
                 shape = "Level",
                 shapekey = shapekey,
                 lab = NULL, # Otherwise will try to label each data point!
                 colby = "Corpus",
                 legendPosition = "none",
                 margingaps = unit(c(0.2, 0.2, 0.8, 0.2), "cm"),
                 colkey = colkey)
dev.off()

```


```{r PCAtools-biplots-register, fig.height=8}
# Biplots to examine components more carefully

# These settings (with legendPosition = "top") were used to generate the legend for the scatterplot matrix above:
#png(here("plots", "PCA_3Ref_Biplot_PC1_PC2test.png"), width = 40, height= 25, units = "cm", res = 300) 
PCAtools::biplot(p,
                 x = "PC1",
                 y = "PC2",
                 lab = NULL, # Otherwise will try to label each data point!
                 colby = "Corpus",
                 pointSize = 1.3,
                 colkey = colkey,
                 shape = "Level",
                 shapekey = shapekey,
                 xlim = c(min(p$rotated[, "PC1"]), max(p$rotated[, "PC1"])),
                 ylim = c(min(p$rotated[, "PC2"]), max(p$rotated[, "PC2"])),
                 showLoadings = FALSE,
                 ellipse = TRUE,
                 axisLabSize = 18,
                 legendPosition = 'right',
                 legendTitleSize = 18,
                 legendLabSize = 14, 
                 legendIconSize = 5) +
  theme(plot.margin = unit(c(0,0,0,0.2), "cm"))

# These settings were used to save this biplot:
#ggsave(here("Plots", "PCA_Ref3TxB_BiplotPC1_PC2.svg"), width = 12, height = 8)

# Biplots to examine components more carefully
PCAtools::biplot(p,
                 x = "PC3",
                 y = "PC4",
                 lab = NULL, # Otherwise will try to label each data point!
                 colby = "Corpus",
                 pointSize = 1.2,
                 colkey = colkey,
                 shape = "Level",
                 shapekey = shapekey,
                 xlim = c(min(p$rotated[, "PC3"]), max(p$rotated[, "PC3"])),
                 ylim = c(min(p$rotated[, "PC4"]), max(p$rotated[, "PC4"])),
                 showLoadings = FALSE,
                 ellipse = TRUE,
                 axisLabSize = 18,
                 legendPosition = 'right',
                 legendTitleSize = 18,
                 legendLabSize = 14, 
                 legendIconSize = 5) +
  theme(plot.margin = unit(c(0,0,0,0.2), "cm"))

#ggsave(here("Plots", "PCA_Ref3TxB_BiplotPC3_PC4.svg"), width = 12, height = 8)

# Biplots to examine components more carefully
PCAtools::biplot(p,
                 x = "PC5",
                 y = "PC6",
                 lab = NULL, # Otherwise will try to label each data point!
                 colby = "Corpus",
                 pointSize = 1.2,
                 colkey = colkey,
                 shape = "Level",
                 shapekey = shapekey,
                 xlim = c(min(p$rotated[, "PC5"]), max(p$rotated[, "PC5"])),
                 ylim = c(min(p$rotated[, "PC6"]), max(p$rotated[, "PC6"])),
                 showLoadings = FALSE,
                 ellipse = TRUE,
                 axisLabSize = 18,
                 legendPosition = 'right',
                 legendTitleSize = 18,
                 legendLabSize = 14, 
                 legendIconSize = 5) +
  theme(plot.margin = unit(c(0,0,0,0.2), "cm"))

#ggsave(here("Plots", "PCA_Ref3TxB_BiplotPC5_PC6.svg"), width = 12, height = 8)

```


```{r PCAtools-biplots-levels, fig.height=8}

# Biplot with ellipses for Level rather than Register
colkey = c(A="#F9B921", B="#A18A33", C="#BD241E", D="#722672", E="#15274D", `Ref. data`= "darkgrey")
shapekey = c(`Spoken BNC2014 Ref.`=16, `Info Teens Ref.`=17, `Youth Fiction Ref.`=15, `Textbook Fiction`=0, `Textbook Conversation`=1, `Textbook Informative`=2)

PCAtools::biplot(p,
                 x = "PC3",
                 y = "PC4",
                 lab = NULL, # Otherwise will try to label each data point!
                 colby = "Level",
                 pointSize = 1.3,
                 colkey = colkey,
                 shape = "Corpus",
                 shapekey = shapekey,
                 xlim = c(min(p$rotated[, "PC3"]), max(p$rotated[, "PC3"])),
                 ylim = c(min(p$rotated[, "PC4"]), max(p$rotated[, "PC4"])),
                 showLoadings = FALSE,
                 ellipse = TRUE,
                 axisLabSize = 18,
                 legendPosition = 'right',
                 legendTitleSize = 18,
                 legendLabSize = 14, 
                 legendIconSize = 5) +
  theme(plot.margin = unit(c(0,0,0,0.2), "cm"))

#ggsave(here("Plots", "PCA_Ref3TxB_BiplotPC3_PC4_levels.svg"), width = 12, height = 8)

# Biplots to examine components more carefully
PCAtools::biplot(p,
                 x = "PC5",
                 y = "PC6",
                 lab = NULL, # Otherwise will try to label each data point!
                 colby = "Level",
                 pointSize = 1.3,
                 colkey = colkey,
                 shape = "Corpus",
                 shapekey = shapekey,
                 xlim = c(min(p$rotated[, "PC5"]), max(p$rotated[, "PC5"])),
                 ylim = c(min(p$rotated[, "PC6"]), max(p$rotated[, "PC6"])),
                 showLoadings = FALSE,
                 ellipse = TRUE,
                 axisLabSize = 18,
                 legendPosition = 'right',
                 legendTitleSize = 18,
                 legendLabSize = 14, 
                 legendIconSize = 5) +
  theme(plot.margin = unit(c(0,0,0,0.2), "cm"))

#ggsave(here("Plots", "PCA_Ref3TxB_BiplotPC5_PC6_levels.svg"), width = 12, height = 8)
```

## Feature contributions (loadings) on each component
```{r loadings-table}

#data <- readRDS(here("FullMDA", "dataforPCA.rds")) 
colnames(data)
pca <- prcomp(data[,9:ncol(data)], scale.=FALSE) # All quantitative variables

# The rotated data that represents the observations / samples is stored in rotated, while the variable loadings are stored in loadings
loadings <- as.data.frame(pca$rotation[,1:4])


# Function that applies a cut-off point for "relevant" loadings (not used in thesis)
smallToZero <- function(x) {if_else(x < 0.05 & x > -0.05, 0, x)}
loadings %>% 
  filter_all(any_vars(. > abs(0.05))) %>% 
  mutate_all(smallToZero) %>% 
  round(3)

# Tabular overiew of loadings with no threshold
loadings %>% 
  round(2)

#write_last_clip()
```

## Graphs of variables
```{r graphs-of-variables, fig.height=16, fig.width=16}

factoextra::fviz_pca_var(pca,
             axes = c(1,2),
             select.var = list(contrib = 30),
             col.var = "contrib", # Colour by contributions to the PC
             gradient.cols = c("#F9B921", "#DB241E", "#722672"),
             title = "",
             repel = TRUE, # Try to avoid too much text overlapping
             ggtheme = ggthemes::theme_few())
#ggsave(here("Plots", "fviz_pca_var_PC1_PC2_Ref3Reg.svg"), width = 11, height = 9)

factoextra::fviz_pca_var(pca,
             axes = c(3,2),
             select.var = list(contrib = 30),
             col.var = "contrib", # Colour by contributions to the PC
             gradient.cols = c("#F9B921", "#DB241E", "#722672"),
             title = "",
             repel = TRUE, # Try to avoid too much text overlapping
             ggtheme = ggthemes::theme_few())
#ggsave(here("Plots", "fviz_pca_var_PC3_PC2_Ref3Reg.svg"), width = 9, height = 8)

factoextra::fviz_pca_var(pca,
             axes = c(3,4),
             select.var = list(contrib = 30),
             col.var = "contrib", # Colour by contributions to the PC
             gradient.cols = c("#F9B921", "#DB241E", "#722672"),
             title = "",
             repel = TRUE, # Try to avoid too much text overlapping
             ggtheme = ggthemes::theme_few())
#ggsave(here("Plots", "fviz_pca_var_PC3_PC4_Ref3Reg.svg"), width = 9, height = 8)

factoextra::fviz_pca_var(pca,
             axes = c(5,6),
             select.var = list(contrib = 25),
             col.var = "contrib", # Colour by contributions to the PC
             gradient.cols = c("#F9B921", "#DB241E", "#722672"),
             title = "",
             repel = TRUE, # Try to avoid too much text overlapping
             ggtheme = ggthemes::theme_few())
#ggsave(here("Plots", "fviz_pca_var_PC5_PC6_Ref3Reg.svg"), width = 9, height = 8)

```

## Exploring feature contributions
```{r feature-contributions, fig.height=30, fig.width=20}

library(gridExtra)

ncounts <- readRDS(here("FullMDA", "counts3Reg.rds"))

ncounts %>%
  group_by(Register, Level) %>% 
  summarise(median(Words), mad(Words))

ncounts %>% 
  filter(Register=="Informative") %>% 
  #filter(Level %in% c("C", "D", "E")) %>% 
  select(Level, VBD, PEAS) %>% 
  group_by(Level) %>% 
  summarise_if(is.numeric, median) %>% 
  mutate_if(is.numeric, round, 2)

cols = c("#F9B921", "#A18A33", "#BD241E", "#722672", "#15274D", "darkgrey")

boxfeature <- ncounts %>% 
  filter(Register=="Informative") %>% 
  #filter(Level %in% c("C", "D", "E")) %>% 
  select(Level, FPP1S, SPP2, CONT, EXIST, AWL, XX0, PASS, VBN) %>% 
  ggplot(aes(x = Level, y = CONT, colour = Level, fill = Level)) +
  geom_jitter(size=0.7, alpha=.7) +
  geom_boxplot(outlier.shape = NA, fatten = 2, fill = "white", alpha = 0.3) +
  scale_colour_manual(values = cols) +
  theme_minimal() +
  theme(legend.position = "none") +
  xlab("")

SPP2 <- boxfeature + aes(y = SPP2) + ylim(c(0,80))# These y-axis limits remove individual outliers that overextend the scales and make the real differences invisible
EXIST <- boxfeature + aes(y = EXIST) + ylim(c(0,25)) 
FFP1 <- boxfeature + aes(y = FPP1S) + ylim(c(0,60))  #
AWL <- boxfeature + aes(y = AWL)
XX0 <- boxfeature + aes(y = XX0) + ylim(c(0,25))
PASS <- boxfeature + aes(y = PASS)
VBN <- boxfeature + aes(y = VBN) + ylim(c(0,40))

boxplots <- grid.arrange( PASS, VBN, AWL, EXIST, FFP1, SPP2, boxfeature, XX0, ncol=2, nrow=4)

#ggsave(here("Plots", "BoxplotsInformativeFeatures.svg"), plot = boxplots, dpi = 300, width = 9, height = 11)

```

## Descriptive PCA results
```{r PCA-results}
## PCA
#data <- readRDS(here("FullMDA", "dataforPCA.rds")) 
#colnames(data)
pca <- prcomp(data[,9:ncol(data)], scale.=FALSE) # All quantitative variables

## Access to the PCA results
#colnames(data)
res.ind <- cbind(data[,1:8], as.data.frame(pca$x)[,1:4])
head(res.ind)

## Examine individual text values
res.ind %>% 
  filter(PC1 < 0.5 & PC1 > -1.5 & PC2 < 1.5 & PC2 > -1.5 & PC3 > 0.5) %>% 
  filter(Subcorpus == "Textbook Informative") %>% 
  select(Filename, Level)

res.ind %>% 
  filter(PC4 < -1.5) %>% 
  filter(Register == "Conversation") %>% 
  select(Filename, Level)

## Summary statistics
res.ind %>% 
  group_by(Register, Corpus) %>% 
  summarise_if(is.numeric, c(mean = mean, sd = sd)) %>% 
  as.data.frame()  

res.ind %>% 
  group_by(Subcorpus, Level) %>% 
  summarise_if(is.numeric, c(mean = mean, sd = sd)) %>% 
  as.data.frame()

library(gtsummary)
res.ind <- res.ind %>% 
  mutate(Subsubcorpus = paste(Corpus, Register, Level, sep = "_")) %>% 
  mutate(Subsubcorpus = as.factor(Subsubcorpus))
  
res.ind %>% 
  select(PC1, PC2, PC3, PC4, Subsubcorpus) %>% 
  tbl_summary(by = Subsubcorpus,
              digits = list(all_continuous() ~ c(2, 2)),
              statistic = all_continuous() ~  "{mean} ({sd})")
  
res.ind %>% 
  select(Register, Level, PC4) %>% 
  group_by(Register, Level) %>% 
  summarise_if(is.numeric, c(Median = median, MAD = mad))

# Search for example texts to illustrate results
res.ind %>% 
  filter(PC3 > 2.5 & PC2 < -2) %>% 
  #filter(Register=="Conversation") %>% 
  #filter(Level == "B") %>% 
  #filter(PC1 > 4.7) %>% 
  select(Filename, PC1, PC2, PC3) %>% 
  mutate(across(where(is.numeric), round, 2))

```

## Raincloud plots
```{r rainplots}

res.ind$Subcorpus <- fct_relevel(res.ind$Subcorpus, "Spoken BNC2014 Ref.", "Textbook Conversation", "Youth Fiction Ref.", "Textbook Fiction", "Info Teens Ref.", "Textbook Informative")

colours <- suf_palette(name = "london", n = 6, type = "continuous") 
colours2 <- suf_palette(name = "classic", n = 5, type = "continuous") 
colours <- c(colours, colours2[c(2:4)]) # Nine colours range
palette <- colours[c(1,5,6,2,3,8,7,4,9)] # Good order for PCA

colours <- palette[c(1,8,9,2,7,3)]

ggplot(res.ind, aes(x=Subcorpus,y=PC1, fill = Subcorpus, colour = Subcorpus))+ # Or leave out "colour = Register" to keep the dots in black
  geom_flat_violin(position = position_nudge(x = .25, y = 0),adjust = 2, trim = FALSE)+
  geom_point(position = position_jitter(width = .15), size = .25)+
# note that here we need to set the x-variable to a numeric variable and bump it to get the boxplots to line up with the rainclouds. 
  geom_boxplot(aes(x = as.numeric(Subcorpus)+0.25, y = PC1), outlier.shape = NA, alpha = 0.3, width = .15, colour = "BLACK") +
  ylab('PC1')+ 
  theme_cowplot()+
  theme(axis.title.x=element_blank())+
  guides(fill = "none", colour = "none") +
  scale_colour_manual(values = colours)+
  scale_fill_manual(values = colours) +
  annotate(geom = "text", x = 1.5, y = -7, label = "Conversation", size = 5) +
  annotate(geom = "segment", x = 0.7, xend = 2.5, y = -6.5, yend = -6.5) +
  annotate(geom = "text", x = 3.5, y = -7, label = "Fiction", size = 5) +
  annotate(geom = "segment", x = 2.7, xend = 4.5, y = -6.5, yend = -6.5) +
  annotate(geom = "text", x = 5.7, y = -7, label = "Informative", size = 5) +
  annotate(geom = "segment", x = 4.7, xend = 6.5, y = -6.5, yend = -6.5) +
  scale_x_discrete(labels=rep(c("Reference", "Textbook"), 3))+
  scale_y_continuous(sec.axis = dup_axis(name=NULL), breaks = seq(from = -6, to = 5, by = 1))

#ggsave(here("Plots", "PC11_3RegComparison.svg"), width = 13, height = 8)
#ggsave(here("Plots", "PC1_3RegComparison.png"), width = 20, height = 15, units = "cm", dpi = 300)

```

## Mixed-effects models predicting PC scores

### Adding the 'source' variable for the random effects
```{r source-variable}

library(stringr)

# Add Source variable for random effect variable in mixed effect models
res.ind <- res.ind %>% 
  mutate(Source = case_when(
  Corpus=="Youth.Fiction" ~ paste("Book", str_extract(Filename, "[0-9]{1,3}"), sep = ""),
  Corpus=="Spoken.BNC2014" ~ "Spoken.BNC2014",
  Corpus=="Textbook.English" ~ as.character(Series),
  Corpus=="Informative.Teens" ~ str_extract(Filename, "BBC|Science_Tech"),
  TRUE ~ "NA")) %>% 
  mutate(Source = case_when(
  Corpus=="Informative.Teens" & is.na(Source) ~ str_remove(Filename, "_.*"),
  TRUE ~ as.character(Source))) %>% 
  mutate(Source = as.factor(Source)) %>% 
  mutate(Corpus = case_when(
    Corpus=="Textbook.English" ~ "Textbook",
    Corpus=="Informative.Teens" ~ "Reference",
    Corpus=="Spoken.BNC2014" ~ "Reference",
    Corpus=="Youth.Fiction" ~ "Reference"
  )) %>% 
  mutate(Corpus = as.factor(Corpus))

summary(res.ind$Source)

# Change the reference level to a theoretically more meaningful level and one that is better populated (see, e.g., https://stats.stackexchange.com/questions/430770/in-a-multilevel-linear-regression-how-does-the-reference-level-affect-other-lev)
summary(res.ind$Corpus)
summary(res.ind$Subcorpus)
summary(res.ind$Level)

res.ind$Level <- relevel(res.ind$Level, "Ref.")
res.ind$Subcorpus <- factor(res.ind$Subcorpus, levels = c("Spoken BNC2014 Ref.", "Textbook Conversation", "Youth Fiction Ref.", "Textbook Fiction", "Info Teens Ref.", "Textbook Informative"))
res.ind$Corpus <- relevel(res.ind$Corpus, "Reference")
```

## PC1
```{r testing-PCA-results-PC1}
# Mixed-effects models
# PC1

# no random effects
summary(lm(PC1 ~ Level:Register, data = res.ind))

# with random effects (slopes + intercepts)
md_source <- lmer(PC1 ~ 1 + (Register|Source), res.ind, REML = FALSE) 
md_corpus <- update(md_source, .~. + Level)
md_register <- update(md_corpus, . ~ . + Register)
md_both <- update(md_corpus, .~. + Register)
md_interaction <- update(md_both, . ~ . + Level:Register)

anova(md_source, md_corpus, md_register, md_both, md_interaction)

tab_model(md_interaction, wrap.labels = 200)

# Random effects
ranef <- as.data.frame(ranef(md_interaction))

ranef %>% 
  filter(grp == "TeenVogue")

```

### Testing assumptions

This chunk was used to check the assumptions of all the other models below.
```{r lmer-diagnostics}

# cf. https://stackoverflow.com/questions/63751541/plot-does-not-show-all-diagnostic-plots-for-lme-lmer

# check distribution of residuals
plot(md_interaction)

# scale-location plot
plot(md_interaction,
     sqrt(abs(resid(.)))~fitted(.),
     type=c("p","smooth"), col.line=1)

# Q-Q plot
lattice::qqmath(md_interaction)

```

## PC2 to PC6
```{r testing-PCA-results-PC2-PC6}
# PC2
# Note that the model with random intercepts and slopes failed to converge for PC2 which is why only slopes are modelled here
md_source <- lmer(PC2 ~ 1 + (1|Source), res.ind, REML = FALSE) 
md_corpus <- update(md_source, .~. + Level)
md_register <- update(md_corpus, . ~ . + Register)
md_both <- update(md_corpus, .~. + Register)
md_interaction <- update(md_both, . ~ . + Level:Register)

anova(md_source, md_corpus, md_register, md_both, md_interaction)

tab_model(md_interaction, wrap.labels = 200)

# PC3
#md_source <- lmer(PC3 ~ 1 + (Register|Source), res.ind, REML = FALSE) # Fails to converge
md_source <- lmer(PC3 ~ 1 + (1|Source), res.ind, REML = FALSE)
md_corpus <- update(md_source, .~. + Level)
md_register <- update(md_corpus, . ~ . + Register)
md_both <- update(md_corpus, .~. + Register)
md_interaction <- update(md_both, . ~ . + Level:Register)

anova(md_source, md_corpus, md_register, md_both, md_interaction)

tab_model(md_interaction, wrap.labels = 200)

# PC4
#md_source <- lmer(PC4 ~ 1 + (Register|Source), res.ind, REML = FALSE) # Singular fit
md_source <- lmer(PC4 ~ 1 + (1|Source), res.ind, REML = FALSE)
md_corpus <- update(md_source, .~. + Level)
md_register <- update(md_corpus, . ~ . + Register)
md_both <- update(md_corpus, .~. + Register)
md_interaction <- update(md_both, . ~ . + Level:Register)

anova(md_source, md_corpus, md_register, md_both, md_interaction)

tab_model(md_interaction, wrap.labels = 200)

```

## PC5 and PC6 were not included in the PhD thesis in the end.

```{r eval=FALSE}

# PC5
#md_source <- lmer(PC5 ~ 1 + (Register|Source), res.ind, REML = FALSE) # Fails to converge
md_source <- lmer(PC5 ~ 1 + (1|Source), res.ind, REML = FALSE) 
md_corpus <- update(md_source, .~. + Level)
md_register <- update(md_corpus, . ~ . + Register)
md_both <- update(md_corpus, .~. + Register)
md_interaction <- update(md_both, . ~ . + Level:Register)

anova(md_source, md_corpus, md_register, md_both, md_interaction)

tab_model(md_interaction, wrap.labels = 200)

# PC6
#md_source <- lmer(PC6 ~ 1 + (Register|Source), res.ind, REML = FALSE) # Fails to converge
md_source <- lmer(PC6 ~ 1 + (1|Source), res.ind, REML = FALSE) 
md_corpus <- update(md_source, .~. + Level)
md_register <- update(md_corpus, . ~ . + Register)
md_both <- update(md_corpus, .~. + Register)
md_interaction <- update(md_both, . ~ . + Level:Register)

anova(md_source, md_corpus, md_register, md_both, md_interaction)

tab_model(md_interaction, wrap.labels = 200)

```

This chunk was used to generate plots for all of the models computed in the chunk above.

```{r PCA-results-lmer-plots, fig.height=15}
# Tweak plot aesthetics with: https://cran.r-project.org/web/packages/sjPlot/vignettes/custplot.html
# Colour customisation trick from: https://stackoverflow.com/questions/55598920/different-line-colors-in-forest-plot-output-from-sjplot-r-package

plot_model(md_interaction, 
           #type = "re", # Option to visualise random effects 
           show.intercept = TRUE,
           show.values=TRUE, 
           show.p=TRUE,
           value.offset = .4,
           value.size = 3.5,
           colors = palette[c(1:3,7:9)],
           group.terms = c(1,5,5,5,5,5,6,4,2,2,2,2,2,3,3,3,3,3), 
           title="Fixed effects",
           wrap.labels = 40,
           axis.title = "PC6 estimated coefficients") +
  theme_sjplot2() 

#ggsave(here("Plots", "TxBRef3Reg_PCA6_lmer_fixed.svg"), height = 6, width = 7)

#svg(here("Plots", "TxBReg3Reg_predicted_PC6_scores_interactions.svg"), height = 8, width = 9)
visreg(md_interaction, xvar = "Level", by="Register", 
       #type = "contrast",
       type = "conditional",
       line=list(col="darkred"), 
       points=list(cex=0.3),
       xlab = "Ref. corpora and textbook level (A to E)", ylab = "PC6",
       layout=c(3,1)
)
#dev.off()

# Random effects
ranef <- as.data.frame(ranef(md_interaction))

ranef %>% 
  filter(grp %in% c("TeenVogue", "BBC", "Dogo", "Ducksters", "Encyclopedia", "Factmonster", "History", "Quatr", "Revision", "Science", "Science_Tech", "Teen", "TweenTribute", "WhyFiles", "World")) %>% 
  ggplot(aes(x = grp, y = condval)) +
  geom_point() +
  coord_flip()
  
ranef %>% 
  filter(grp %in% levels(data$Series)) %>% 
  ggplot(aes(x = grp, y = condval)) +
  geom_point() +
  coord_flip()

# Modeling subcorpora
md_subcorpus <- lmer(PC1 ~ Subcorpus + (1|Source), res.ind, REML = FALSE)
tab_model(md_subcorpus)

plot_model(md_subcorpus, 
           #type = "re", # Option to visualise random effects 
           show.intercept = TRUE,
           show.values=TRUE, 
           show.p=TRUE,
           value.offset = .4,
           value.size = 3.5,
           colors = palette[c(1:3,7:9)],
           group.terms = c(1,5,2,6,4,3), 
           wrap.labels = 40,
           axis.title = "PC6 estimated coefficients",
           title="Fixed effects") +
  theme_sjplot2()

subcorpus_results <- emmeans(md_subcorpus, "Subcorpus")
summary(subcorpus_results)
#write_last_clip()

```


# Package used in this script
```{r package-citations}

#packages.bib <- sapply(1:length(loadedNamespaces()), function(i) toBibtex(citation(loadedNamespaces()[i])))

knitr::write_bib(c(.packages(), "knitr"), "packages.bib")

sessionInfo()

```
