---
title: "A new model of intra-textbook variation"
author: "Elen Le Foll"
date: "09/11/2021"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
  editor_options: 
    markdown: 
      wrap: sentence

---

## Set-up

*Built with R `r getRversion()`*  
*Last saved on `r format(Sys.Date(), "%d %B %Y")` at `r format(Sys.time(), "%H:%M")`*

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)

#library(car) # For data wrangling
library(caret) # For its confusion matrix function
#library(clipr) # For easy exports
library(cowplot)
library(DescTools) # For 95% CI
library(dplyr)
library(emmeans)
#library(FactoMineR) # For Shiny app
library(factoextra) # For circular graphs of variables
library(here) # For dynamic file paths
library(ggplot2) 
library(ggthemes) # For theme of factoextra plots
library(lme4) # For linear regression modelling
library(patchwork) # To create figures with more than one plot
library(pca3d) # For 3-D plots
library(PCAtools) # For nice biplots of PCA results
#library(purrr) # For data wrangling
library(psych) # For various useful stats function
library(sjPlot) # For model plots and tables
library(suffrager) # For pretty feminist colour palettes :)
library(tidyr)
library(tibble)
library(visreg) # For plots of interaction effects

source(here("R_rainclouds.R")) # For geom_flat_violin rainplots

```


```{r colours, include=FALSE}
#library(suffrager)
# suf_palette(name, n, type = c("discrete", "continuous"))

colours <- suf_palette(name = "london", n = 6, type = "continuous") 
colours2 <- suf_palette(name = "classic", n = 5, type = "continuous") 
colours <- c(colours, colours2[c(2:4)]) # Nine colours range
palette <- colours[c(1,5,6,2,3,8,7,4,9)] # Good order for PCA
scales::show_col(palette)

```


# Performing the PCA

## TEC data import

```{r quick-TEC-import}

TxBcounts <- readRDS(here("FullMDA", "TxBcounts3.rds"))
colnames(TxBcounts)
nrow(TxBcounts)

TxBzlogcounts <- readRDS(here("FullMDA", "TxBzlogcounts.rds")) 
nrow(TxBzlogcounts)
colnames(TxBzlogcounts)

TxBdata <- cbind(TxBcounts[,1:6], as.data.frame(TxBzlogcounts))
str(TxBdata)
ncol(TxBdata)-6

#saveRDS(TxBdata, here("FullMDA", "TxBdata.rds")) # Last saved 6 December 2021

```

## Metadata on TEC

```{r TEC-metadata}

TxBdata %>% summarise(total = sum(Words))

metadata <- TxBdata %>% 
  select(Filename, Country, Series, Level, Register, Words) %>% 
  mutate(Volume = paste(Series, Level)) %>% 
  mutate(Volume = fct_rev(Volume)) %>% 
  mutate(Volume = fct_reorder(Volume, as.numeric(Level))) %>% 
  group_by(Volume) %>% 
  mutate(wordcount = sum(Words)) %>% 
  ungroup() %>% 
  distinct(Volume, .keep_all = TRUE)

metadata

ggplot(metadata, aes(x = Volume, y = wordcount, fill = Country)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = palette[c(3,8,1)]) +
  scale_y_continuous(limits = c(0,90000), expand = c(0,0), labels = scales::comma) +
  coord_flip() +
  theme_bw() +
  labs(x = "", y = "Number of words") +
  theme(legend.position = c(0.73,0.1))

#ggsave(here("plots", "TEC-T_wordcounts.svg"), width = 6, height = 10)

metadataInstr <- TxBdata %>% 
  select(Country, Series, Level, Register, Words) %>% 
  filter(Register=="Instructional") %>% 
  mutate(Volume = paste(Series, Register)) %>% 
  mutate(Volume = fct_rev(Volume)) %>% 
  mutate(Volume = fct_reorder(Volume, as.numeric(Level))) %>% 
  group_by(Volume, Register) %>% 
  mutate(InstrWordcount = sum(Words)) %>% 
  ungroup() %>% 
  distinct(Volume, .keep_all = TRUE) %>% 
  select(Series, InstrWordcount)

metadataInstr

metaWordcount <- TxBdata %>% 
  select(Country, Series, Level, Register, Words) %>% 
  group_by(Series) %>% 
  mutate(TECwordcount = sum(Words)) %>% 
  ungroup() %>% 
  distinct(Series, .keep_all = TRUE) %>% 
  select(Series, TECwordcount)

wordcount <- merge(metaWordcount, metadataInstr, by = "Series")

wordcount %>% 
  mutate(InstrucPercent = InstrWordcount/TECwordcount*100) %>% 
  arrange(InstrucPercent) %>% 
  mutate(InstrucPercent = round(InstrucPercent, 2))

```


## Checking the factorability of data

```{r factorability.TEC}

colnames(TxBdata)
ncol(TxBdata) - 6
kmo <- KMO(TxBdata[,7:ncol(TxBdata)]) 
kmo # 0.86 (if conducted on counts all normalised per 100 words, this value drops to 0.61!)
kmo$MSAi[order(kmo$MSAi)]

# Remove first feature with MSAs of < 0.5
TxBdata <- TxBdata %>% 
  select(-c(MDWO))

kmo <- KMO(TxBdata[,7:ncol(TxBdata)]) 
kmo # 0.87
kmo$MSAi[order(kmo$MSAi)] # All individual MSA > 0.5

```

## Choosing the number of principal components to retain and excluding features with low final communalites

https://rdrr.io/cran/FactorAssumptions/man/communalities_optimal_solution.html

```{r screeplot-communalities}

#TxBdata <- readRDS(here("FullMDA", "TxBdata.rds")) %>% select(-MDWO)
#colnames(TxBdata)

# Plot screen plot
#png(here("plots", "screeplot-TEC-only.png"), width = 20, height= 12, units = "cm", res = 300)
scree(TxBdata[,7:ncol(TxBdata)], factors = FALSE, pc = TRUE) # Retain six components
dev.off()

# Perform PCA
pca1 <- psych::principal(TxBdata[,7:ncol(TxBdata)], 
                         nfactors = 6)
pca1$loadings

pca1$communality %>% sort(.) # If features with communalities of < 0.2 are removed, we remove STPR, HDG, MDNE and CAUSE

TxBdataforPCA <- TxBdata %>% 
  select(-c(STPR, MDNE, HDG, CAUSE))

kmo <- KMO(TxBdataforPCA[,7:ncol(TxBdataforPCA)]) 
kmo # 0.88
kmo$MSAi[order(kmo$MSAi)] # All individual MSA > 0.5

# Final number of features
ncol(TxBdataforPCA)-6

#saveRDS(TxBdataforPCA, here("FullMDA", "TxBdataforPCA.rds")) # Last saved on 3 December 2021

```

https://stats.stackexchange.com/questions/59213/how-to-compute-varimax-rotated-principal-components-in-r

### Testing the effect of rotating the components
This chunk was used when considering whether or not to rotate the components (see methods section)
```{r PCA-rotation-tests, eval=FALSE}

# Comparing a rotated vs. a non-rotated solution

TxBdata <- readRDS(here("FullMDA", "TxBdataforPCA.rds"))
colnames(TxBdata)
ncol(TxBdata)-6

# No rotation
pca2 <- psych::principal(TxBdata[,7:ncol(TxBdata)], 
                         nfactors = 6, 
                         rotate = "none")

pca2$loadings

biplot.psych(pca2, 
             vars = TRUE, 
             choose=c(1,2),
             )

# Promax rotation
pca2.rotated <- psych::principal(TxBdata[,7:ncol(TxBdata)], 
                         nfactors = 6, 
                         rotate = "promax")

# This summary shows the component correlations which is particularly interesting
pca2.rotated

pca2.rotated$loadings

biplot.psych(pca2.rotated, vars = TRUE, choose=c(1,2))

```

## Peforming the PCA

### Using the full data (except outliers)
```{r PCA-prcomp-full-data}

# Perform PCA on full data
TxBdata <- readRDS(here("FullMDA", "TxBdataforPCA.rds"))
nrow(TxBdata)
ncol(TxBdata)-6
```

#### Using subsets of the data

Perform PCA on random subset of the data to test the stability of the solution. Re-running this line will generate a new subset of the TEC texts containing 2/3 of the texts randomly sampled.

```{r PCA-prcomp-random-subset, eval=FALSE}

TxBdata <- readRDS(here("FullMDA", "TxBdataforPCA.rds")) %>% 
  slice_sample(n = 1961*0.6, replace = FALSE)
nrow(TxBdata)
TxBdata$Filename[1:10]

colnames(TxBdata)
nrow(TxBdata) / (ncol(TxBdata)-6) # Check that there is enough data to conduct a PCA. This should be > 5 (see Friginal & Hardy 2014: 303–304).
```

Perform PCA on country subset of the data to test the stability of the solution:

```{r PCA-prcomp-country-subset, eval=FALSE}

TxBdata <- readRDS(here("FullMDA", "TxBdataforPCA.rds")) %>% 
  #filter(Country == "France")
  #filter(Country == "Germany")
  filter(Country == "Spain")

nrow(TxBdata)
TxBdata$Filename[1:10] # Check data

colnames(TxBdata)
nrow(TxBdata) / (ncol(TxBdata)-6) # Check that there is enough data to conduct a PCA. This should be > 5 (see Friginal & Hardy 2014: 303–304).

```

### PCA code

```{r PCA-prcomp}

pca <- prcomp(TxBdata[,7:ncol(TxBdata)], scale.=FALSE) # All quantitative variables for all TxB files except outliers
register  <- factor(TxBdata[,"Register"]) # Register
level <- factor(TxBdata[,"Level"]) # Textbook proficiency level

summary(register)
summary(level)
summary(pca)

scree(TxBdata[,7:ncol(TxBdata)], factors = FALSE, pc = TRUE) # The scree plots pretty much always suggest that six components should be retained

```

## Plotting PCA results 

### 3D plots
```{r 3D-plots-TxB, eval=FALSE}

## 3-D PCA plots https://cran.r-project.org/web/packages/pca3d/vignettes/pca3d.pdf ##
# 3-D plot
colours <- palette[c(1:3,8,7)] # without poetry
names(colours) <- c("Conversation", "Fiction", "Informative", "Instructional", "Personal")
scales::show_col(colours) # Check colours
pca3d(pca, 
      group = register,
      components = 1:3,
      #components = 4:6,
      show.ellipses=FALSE, 
      ellipse.ci=0.75,
      show.plane=FALSE,
      col = col,
      shape = "sphere",
      radius = 1,
      legend = "right")

snapshotPCA3d(here("plots", "PCA_TxB_3Dsnapshot.png"))

names(colours) <- c("C", "B", "E", "A", "D") # To colour the dots according to the profiency level of the textbooks
pca3d(pca, 
      components = 4:6,
      group = level,
      show.ellipses=FALSE, 
      ellipse.ci=0.75,
      show.plane=FALSE,
      col = colours,
      shape = "sphere",
      radius = 0.8,
      legend = "right")

```

### Bi-plots
```{r PCAtools-biplots-TxB, warning=FALSE, fig.height=15, fig.width=20}
# PCA for PCAtools

# This package requires the data to be formatted in a rather unconventional way so it needs to wrangled first.

#TxBdata <- readRDS(here("FullMDA", "TxBdataforPCA.rds"))

TxBdata2meta <- TxBdata[,1:6]
rownames(TxBdata2meta) <- TxBdata2meta$Filename
TxBdata2meta <- TxBdata2meta %>% select(-Filename)
head(TxBdata2meta)

TxBdata2 = TxBdata
rownames(TxBdata2) <- TxBdata2$Filename
TxBdata2num <- as.data.frame(base::t(TxBdata2[,7:ncol(TxBdata2)]))
TxBdata2num[1:12,1:3] # Check sanity of data

p <- PCAtools::pca(TxBdata2num, 
         metadata = TxBdata2meta,
         scale = FALSE)

colkey = c(Conversation="#BD241E", Fiction="#A18A33", Informative="#15274D", Instructional="#F9B921", Personal="#722672")
shapekey = c(A=1, B=2, C=6, D=0, E=5)

# Biplots to examine components more carefully

#png(here("plots", "PCA_TxB_Biplot_PC1_PC2.png"), width = 40, height= 25, units = "cm", res = 300)
PCAtools::biplot(p,
                 x = "PC1",
                 y = "PC2",
                 lab = NULL, # Otherwise will try to label each data point!
                 #xlim = c(min(p$rotated$PC1)-0.5, max(p$rotated$PC1)+0.5),
                 #ylim = c(min(p$rotated$PC2)-0.5, max(p$rotated$PC2)+0.5),
                 colby = "Register",
                 pointSize = 2,
                 colkey = colkey,
                 shape = "Level",
                 shapekey = shapekey,
                 showLoadings = FALSE,
                 ellipse = TRUE,
                 axisLabSize = 22,
                 legendPosition = 'right',
                 legendTitleSize = 22,
                 legendLabSize = 18, 
                 legendIconSize = 7) +
  theme(plot.margin = unit(c(0,0,0,0.2), "cm"))
#dev.off()
#ggsave(here("plots", "PCA_TxB_BiplotPC1_PC2.svg"), width = 12, height = 10)

# Biplots to examine components more carefully
pRegisters <- PCAtools::biplot(p,
                 x = "PC3",
                 y = "PC4",
                 lab = NULL, # Otherwise will try to label each data point!
                 colby = "Register",
                 pointSize = 2,
                 colkey = colkey,
                 shape = "Level",
                 shapekey = shapekey,
                 showLoadings = FALSE,
                 ellipse = TRUE,
                 legendPosition = 'right',
                 legendTitleSize = 22,
                 legendLabSize = 18, 
                 legendIconSize = 7) +
  theme(plot.margin = unit(c(0,0,0,0.2), "cm"))

#ggsave(here("plots", "PCA_TxB_BiplotPC3_PC4.svg"), width = 12, height = 10)

```


```{r PCAtools-biplots-TxB-Levels, warning=FALSE, fig.height=30, fig.width=20}

# Biplot with ellipses for Level rather than Register
colkey = c(A="#F9B921", B="#A18A33", C="#BD241E", D="#722672", E="#15274D")
shapekey = c(Conversation=1, Fiction=2, Informative=6, Instructional=0, Personal=5)

pLevels <- PCAtools::biplot(p,
                 x = "PC3",
                 y = "PC4",
                 lab = NULL, # Otherwise will try to label each data point!
                 #xlim = c(min(p$rotated$PC1)-0.5, max(p$rotated$PC1)+0.5),
                 #ylim = c(min(p$rotated$PC2)-0.5, max(p$rotated$PC2)+0.5),
                 colby = "Level",
                 pointSize = 2,
                 colkey = colkey,
                 shape = "Register",
                 shapekey = shapekey,
                 showLoadings = FALSE,
                 ellipse = TRUE,
                 legendPosition = 'right',
                 legendTitleSize = 22,
                 legendLabSize = 18, 
                 legendIconSize = 7) +
  theme(plot.margin = unit(c(0,0,0,0.2), "cm"))
#ggsave(here("plots", "PCA_TxB_BiplotPC3_PC4_Level.svg"), width = 12, height = 10)

# Save the two different representations of data points on PC2 and PC3 using the {patchwork} package
pRegisters / pLevels
#ggsave(here("plots", "PCA_TxB_BiplotPC3_PC4_Register_vs_Level.svg"), width = 14, height = 20)

pLevels <- PCAtools::biplot(p,
                 x = "PC5",
                 y = "PC6",
                 lab = NULL, # Otherwise will try to label each data point!
                 xlim = c(min(p$rotated$PC5)-0.5, max(p$rotated$PC5)+0.5),
                 ylim = c(min(p$rotated$PC6)-0.5, max(p$rotated$PC6)+0.5),
                 colby = "Level",
                 pointSize = 2,
                 colkey = colkey,
                 shape = "Register",
                 shapekey = shapekey,
                 showLoadings = FALSE,
                 ellipse = TRUE,
                 #axisLabSize = 22,
                 legendPosition = 'right',
                 legendTitleSize = 22,
                 legendLabSize = 18, 
                 legendIconSize = 7) +
  theme(plot.margin = unit(c(0,0,0,0.2), "cm"))

#pLevels
#ggsave(here("plots", "PCA_TxB_BiplotPC5_PC6_Level.svg"), width = 12, height = 10)

# Biplots to examine components more carefully

colkey = c(Conversation="#BD241E", Fiction="#A18A33", Informative="#15274D", Instructional="#F9B921", Personal="#722672")
shapekey = c(A=1, B=2, C=6, D=0, E=5)

pRegisters <- PCAtools::biplot(p,
                 x = "PC5",
                 y = "PC6",
                 lab = NULL, # Otherwise will try to label each data point!
                 #xlim = c(min(p$rotated$PC1)-0.5, max(p$rotated$PC1)+0.5),
                 #ylim = c(min(p$rotated$PC2)-0.5, max(p$rotated$PC2)+0.5),
                 colby = "Register",
                 pointSize = 2,
                 colkey = colkey,
                 shape = "Level",
                 shapekey = shapekey,
                 showLoadings = FALSE,
                 ellipse = TRUE,
                 axisLabSize = 22,
                 legendPosition = 'right',
                 legendTitleSize = 22,
                 legendLabSize = 18, 
                 legendIconSize = 7) +
  theme(plot.margin = unit(c(0,0,0,0.2), "cm"))
#ggsave("Plots/PCA_TxB_BiplotPC5_PC6.svg", width = 12, height = 10)

# Save the two different representations of data points on PC4 and PC5 using the {patchwork} package
pRegisters / pLevels
#ggsave(here("plots", "PCA_TxB_BiplotPC5_PC6_Register_vs_Level.svg"), width = 14, height = 20)
```


```{r PCAtools-pairsplots-TxB, eval=FALSE}

## Very slow, open in zoomed out window!
# Add legend manually? Yes (take it from the biplot code below), let's not waste too much time, here. Or use Evert's mvar.pairs plot (though that also requires manual axis annotation)

png(here("plots", "PCA_TxB_pairsplot.png"), width = 25, height= 45, units = "cm", res = 300)
PCAtools::pairsplot(p,
                 triangle = FALSE,
                 components = 1:6,
                 ncol = 3,
                 nrow = 5,
                 pointSize = 0.8,
                 lab = NULL, # Otherwise will try to label each data point!
                 colby = "Register",
                 colkey = colkey,
                 shape = "Level",
                 shapekey = shapekey,
                 margingaps = unit(c(0.2, 0.2, 0.8, 0.2), "cm"),
                 legendPosition = "none")
dev.off()

#ggsave(here("plots", "PCA_TxB_pairsplot.svg"), width = 15, height = 15)

```

### Feature contributions (loadings) on each component
```{r graphs-of-variables}

TxBdata <- readRDS(here("FullMDA", "TxBdataforPCA.rds"))

pca <- prcomp(TxBdata[,7:ncol(TxBdata)], scale.=FALSE) # All quantitative variables for all TxB files

# The rotated data that represents the observations / samples is stored in rotated, while the variable loadings are stored in loadings

loadings <- as.data.frame(pca$rotation[,1:4])

loadings %>% round(2)

smallToZero <- function(x) {if_else(x < 0.05 & x > -0.05, 0, x)}
loadings %>% 
  filter_all(any_vars(. > abs(0.05))) %>% 
  mutate_all(smallToZero) %>% 
  round(3)

#write_last_clip()

# Compare frequencies of the individual features across different registers and levels

TxBcounts %>% 
  group_by(Register, Level) %>% 
  summarise(median(NCOMP), MAD(NCOMP)) %>% 
  mutate(across(where(is.numeric), round, 2)) %>% 
  as.data.frame()

# Graph of variables
factoextra::fviz_pca_var(pca,
             axes = c(1,2),
             select.var = list(cos2 = 0.1),
             col.var = "contrib", # Colour by contributions to the PC
             gradient.cols = c("#F9B921", "#DB241E", "#722672"),
             title = "",
             repel = TRUE, # Try to avoid too much text overlapping
             ggtheme = ggthemes::theme_few())
#ggsave(here("plots", "fviz_pca_var_PC1_PC2.svg"), width = 11, height = 9)

factoextra::fviz_pca_var(pca,
             axes = c(3,2),
             select.var = list(contrib = 30),
             col.var = "contrib", # Colour by contributions to the PC
             gradient.cols = c("#F9B921", "#DB241E", "#722672"),
             title = "",
             repel = TRUE, # Try to avoid too much text overlapping
             ggtheme = ggthemes::theme_few())
#ggsave(here("plots", "fviz_pca_var_PC3_PC2.svg"), width = 9, height = 8)

factoextra::fviz_pca_var(pca,
             axes = c(3,4),
             select.var = list(contrib = 30),
             col.var = "contrib", # Colour by contributions to the PC
             gradient.cols = c("#F9B921", "#DB241E", "#722672"),
             title = "",
             repel = TRUE, # Try to avoid too much text overlapping
             ggtheme = ggthemes::theme_few())
#ggsave(here("plots", "fviz_pca_var_PC3_PC4.svg"), width = 9, height = 8)

factoextra::fviz_pca_var(pca,
             axes = c(5,6),
             select.var = list(cos2 = 0.05),
             col.var = "contrib", # Colour by contributions to the PC
             gradient.cols = c("#F9B921", "#DB241E", "#722672"),
             title = "",
             repel = TRUE, # Try to avoid too much text overlapping
             ggtheme = ggthemes::theme_few())
#ggsave(here("plots", "fviz_pca_var_PC5_PC6.svg"), width = 9, height = 8)

```

## Visualising results in shiny app
```{r PCA-FactoMineR, eval=FALSE}

colnames(TxBdata)

res.pca <- FactoMineR::PCA(TxBdata, quali.sup = 1:5, quanti.sup = 6, scale.unit = FALSE, graph = FALSE, ncp = 6)

summary(res.pca, ncp = 6, nbelements = 6)

# For some reason I don't understand, the dimdesc() function only works if I eliminate the supplementary variables (although they obviously don't contribute to the dimensions anyway!)
res.pca2 <- FactoMineR::PCA(TxBdata[,7:ncol(TxBdata)], scale.unit = FALSE, graph = FALSE)
# 
dims <- dimdesc(res.pca2, proba = 0.0001)

dims

dims$Dim.1
dims$Dim.2
dims$Dim.3
dims$Dim.4
dims$Dim.5

plot(res.pca, choix = "ind", habillage = "Register", col.hab = colours[1:5], cex = 1.1, select = "contrib 10", title = "Graphe des individus")

plot(res.pca, choix="var", select="cos2 0.6")  # plot the variables with cos2 greater than 0.6

library(Factoshiny)
TxBdata2 <- TxBdata %>% 
  select(-Words, -Filename)

res.shiny <- PCAshiny(TxBdata2)

```

## Exploring the dimensions of the model
### Descriptive stats of dimensions
```{r PCA-numerical-results}
# http://www.sthda.com/english/articles/31-principal-component-methods-in-r-practical-guide/118-principal-component-analysis-in-r-prcomp-vs-princomp/#pca-results-for-variables

#TxBdata <- readRDS(here("FullMDA", "TxBdataforPCA.rds"))

pca <- prcomp(TxBdata[,7:ncol(TxBdata)], scale.=FALSE) # All quantitative variables for all TxB files
register  <- factor(TxBdata[,"Register"]) # Register
level <- factor(TxBdata[,"Level"]) # Textbook proficiency level

summary(register)
summary(level)
summary(pca)

## Access to the PCA results
pca$rotation[,1]

res.ind <- cbind(TxBdata[,1:5], as.data.frame(pca$x)[,1:6])
head(res.ind)

res.ind %>% 
  group_by(Register) %>% 
  summarise_if(is.numeric, mean)

res.ind %>% 
  group_by(Register, Level) %>% 
  summarise_if(is.numeric, mean) %>% 
  mutate(across(where(is.numeric), round, 2)) %>% 
  as.data.frame()

res.ind %>% 
  select(Register, Level, PC2) %>% 
  group_by(Register, Level) %>% 
  summarise_if(is.numeric, c(Median = median, MAD = mad))%>% 
  mutate(across(where(is.numeric), round, 2)) %>% 
  as.data.frame()

# Search for example texts to illustrate results
res.ind %>% 
  #filter(PC3 > 2.5 & PC2 < -2) %>% 
  filter(PC4 < -2.5) %>% 
  select(Filename, PC3, PC4) %>% 
  mutate(across(where(is.numeric), round, 2))

```

### Mixed-effects models of dimension scores and plots
```{r lmer-models}

# Dimension 1
# Model with only Register as a fixed effect
lm1 <- lm(PC1 ~ Register, data = res.ind)
summary(lm1)

# Models with Textbook series as random intercepts
md1 <- lmer(PC1 ~ Register*Level + (1|Series), data = res.ind, REML = FALSE)
md1Register <- lmer(PC1 ~ Register + (1|Series), data = res.ind, REML = FALSE)
md1Level <- lmer(PC1 ~ Level + (1|Series), data = res.ind, REML = FALSE)

anova(md1, md1Register, md1Level)

tab_model(md1, wrap.labels = 300) # Marginal R2 = 0.890
tab_model(md1Register) # Marginal R2 = 0.877
tab_model(md1Level) # Marginal R2 = 0.002

# Tweak plot aestetics with: https://cran.r-project.org/web/packages/sjPlot/vignettes/custplot.html
# Colour customisation trick from: https://stackoverflow.com/questions/55598920/different-line-colors-in-forest-plot-output-from-sjplot-r-package

# Plot of fixed effects:
plot_model(md1Register, 
           type = "est",
           show.intercept = TRUE,
           show.values=TRUE, 
           show.p=TRUE,
           value.offset = .4,
           value.size = 3.5,
           colors = palette[c(1:3,8,7)],
           group.terms = c(1:5), 
           title = "",
           wrap.labels = 40,
           axis.title = "PC1 estimated coefficients") +
  theme_sjplot2() 

#ggsave(here("plots", "TxB_PCA1_lmer_fixedeffects_Register.svg"), height = 3, width = 8)

Register_results <- emmeans(md1Register, "Register")
summary(Register_results)
comparisons <- pairs(Register_results, adjust = "tukey")
comparisons
#write_last_clip()
confint(comparisons)
#write_last_clip()

TxBdata %>% 
  filter(Register == "Instructional") %>% 
  group_by(Level) %>% 
  summarise(median(VPRT))

# Plot of random effects:
plot_model(md1, 
           type = "re", # Option to visualise random effects
           show.values=TRUE, 
           show.p=TRUE,
           value.offset = .4,
           value.size = 3.5,
           colors = "bw",
           wrap.labels = 40,
           axis.title = "PC1 estimated coefficients") +
  theme_sjplot2()

#ggsave(here("plots", "TxB_PCA1_lmer_randomeffects.svg"), height = 3, width = 8)

# Dimension 2
md2 <- lmer(PC2 ~ Register*Level + (1|Series), data = res.ind, REML = FALSE)
md2Register <- lmer(PC2 ~ Register + (1|Series), data = res.ind, REML = FALSE)
md2Level <- lmer(PC2 ~ Level + (1|Series), data = res.ind, REML = FALSE)
anova(md2, md2Register, md2Level)

tab_model(md2) # Marginal R2 = 0.723
tab_model(md2Register) # Marginal R2 = 0.558
tab_model(md2Level) # Marginal R2 = 0.228

# Plot of fixed effects:
plot_model(md2, 
           type = "est",
           show.intercept = TRUE,
           show.values=TRUE, 
           show.p=TRUE,
           value.offset = .4,
           value.size = 3.5,
           colors = palette[c(1:3,8,7)],
           group.terms = c(1:5,1,1,1,1,2:5,2:5,2:5,2:5), 
           title = "",
           wrap.labels = 40,
           axis.title = "PC2 estimated coefficients") +
  theme_sjplot2() 

#ggsave(here("plots", "TxB_PCA2_lmer_fixedeffects.svg"), height = 8, width = 8)

#svg(here("plots", "TxB_predicted_PC2_scores_interactions.svg"), height = 5, width = 8)
visreg(md2, xvar = "Level", by="Register", type = "conditional",
       line=list(col="darkred"), 
       xlab = "Textbook Level", ylab = "PC2"
       #,gg = TRUE
       ,layout=c(5,1)
)
dev.off()

# Plots of random effects
## Random intercepts
plot_model(md2, 
           type = "re", # Option to visualise random effects
           show.values=TRUE, 
           show.p=TRUE,
           value.offset = .4,
           value.size = 3.5,
           colors = "bw",
           wrap.labels = 40,
           axis.title = "PC2 estimated coefficients") +
  theme_sjplot2()
#ggsave(here("plots", "TxB_PCA2_lmer_randomeffects.svg"), height = 3, width = 8)

# Textbook Series-Register interactions
visreg::visreg(md2, "Register", by="Series", re.form=~(1|Series),
               ylab="PC2", line=list(col="darkred"))

visreg(md2, xvar = "Series", by="Level", type = "conditional", re.form=~(1|Series), 
       line=list(col="darkred"), xlab = "Textbook Series", ylab = "PC2",
       layout=c(1,5))

# Dimension 3
md3 <- lmer(PC3 ~ Register*Level + (1|Series), data = res.ind, REML = FALSE)
md3Register <- lmer(PC3 ~ Register + (1|Series), data = res.ind, REML = FALSE)
md3Level <- lmer(PC3 ~ Level + (1|Series), data = res.ind, REML = FALSE)

anova(md3, md3Register, md3Level)

tab_model(md3) # Marginal R2 = 0.436
tab_model(md3Register) # Marginal R2 = 0.272
tab_model(md3Level) # Marginal R2 = 0.119

# Plot of fixed effects:
plot_model(md3, 
           type = "est",
           show.intercept = TRUE,
           show.values=TRUE, 
           show.p=TRUE,
           value.offset = .4,
           value.size = 3.5,
           colors = palette[c(1:3,8,7)],
           group.terms = c(1:5,1,1,1,1,2:5,2:5,2:5,2:5), 
           title = "",
           wrap.labels = 40,
           axis.title = "PC3 estimated coefficients") +
  theme_sjplot2() 
#ggsave(here("plots", "TxB_PCA3_lmer_fixedeffects.svg"), height = 8, width = 8)

#svg(here("plots", "TxB_predicted_PC3_scores_interactions.svg"), height = 5, width = 8)
visreg(md3, xvar = "Level", by="Register", type = "conditional",
       line=list(col="darkred"), 
       xlab = "Textbook Level", ylab = "PC3"
       #,gg = TRUE
       ,layout=c(5,1)
)
dev.off()

# Plot of random effects:
plot_model(md3, 
           type = "re", # Option to visualise random effects
           show.values=TRUE, 
           show.p=TRUE,
           value.offset = .4,
           value.size = 3.5,
           color = "bw",
           wrap.labels = 40,
           axis.title = "PC3 estimated coefficients") +
  theme_sjplot2()
#ggsave(here("plots", "TxB_PCA3_lmer_randomeffects.svg"), height = 3, width = 8)

# Textbook Series-Register interactions
visreg::visreg(md3, "Series", by="Register", re.form=~(1|Series),
               ylab="PC2", line=list(col="darkred"))

visreg(md2, xvar = "Level", by="Series", type = "conditional", re.form=~(1|Series), 
       line=list(col="darkred"), xlab = "Textbook Series", ylab = "PC2")

# Dimension 4
md4 <- lmer(PC4 ~ Register*Level + (1|Series), data = res.ind, REML = FALSE)
md4Register <- lmer(PC4 ~ Register + (1|Series), data = res.ind, REML = FALSE)
md4Level <- lmer(PC4 ~ Level + (1|Series), data = res.ind, REML = FALSE)

anova(md4, md4Register, md4Level)

tab_model(md4) # Marginal R2 = 0.426
tab_model(md4Register) # Marginal R2 = 0.203
tab_model(md4Level) # Marginal R2 = 0.187

# Plot of fixed effects:
plot_model(md4, 
           type = "est",
           show.intercept = TRUE,
           show.values=TRUE, 
           show.p=TRUE,
           value.offset = .4,
           value.size = 3.5,
           colors = palette[c(1:3,8,7)],
           group.terms = c(1:5,1,1,1,1,2:5,2:5,2:5,2:5), 
           title = "",
           wrap.labels = 40,
           axis.title = "PC4 estimated coefficients") +
  theme_sjplot2() 
#ggsave(here("plots", "TxB_PCA4_lmer_fixedeffects.svg"), height = 8, width = 8)

#svg(here("plots", "TxB_predicted_PC4_scores_interactions.svg"), height = 5, width = 8)
visreg(md4, xvar = "Level", by="Register", type = "conditional",
       line=list(col="darkred"), 
       xlab = "Textbook Level", ylab = "PC4"
       #,gg = TRUE
       ,layout=c(5,1)
)
dev.off()


# Plot of random effects:
plot_model(md4, 
           type = "re", # Option to visualise random effects
           show.values=TRUE, 
           show.p=TRUE,
           value.offset = .4,
           value.size = 3.5,
           color = "bw",
           wrap.labels = 40,
           axis.title = "PC4 estimated coefficients") +
  theme_sjplot2()

#ggsave(here("plots", "TxB_PCA4_lmer_randomeffects.svg"), height = 3, width = 8)

# Dimension 5
md5 <- lmer(PC5 ~ Register*Level + (1|Series), data = res.ind, REML = FALSE)
md5Register <- lmer(PC5 ~ Register + (1|Series), data = res.ind, REML = FALSE)
md5Level <- lmer(PC5 ~ Level + (1|Series), data = res.ind, REML = FALSE)

anova(md5, md5Register, md5Level)

tab_model(md5) # Marginal R2 = 0.071
tab_model(md5Register) # Marginal R2 = 0.001
tab_model(md5Level) # Marginal R2 = 0.055

# Plot of fixed effects:
plot_model(md5, 
           type = "est",
           show.intercept = TRUE,
           show.values=TRUE, 
           show.p=TRUE,
           value.offset = .4,
           value.size = 3.5,
           colors = palette[c(1:3,8,7)],
           group.terms = c(1:5,1,1,1,1,2:5,2:5,2:5,2:5), 
           title = "",
           wrap.labels = 40,
           axis.title = "PC5 estimated coefficients") +
  theme_sjplot2() 

#ggsave(here("plots", "TxB_PCA5_lmer_fixedeffects.svg"), height = 3, width = 8)

# Plot of random effects:
plot_model(md5, 
           type = "re", # Option to visualise random effects
           show.values=TRUE, 
           show.p=TRUE,
           value.offset = .4,
           value.size = 3.5,
           color = "bw",
           wrap.labels = 40,
           axis.title = "PC5 estimated coefficients") +
  theme_sjplot2()

#ggsave(here("plots", "TxB_PCA5_lmer_randomeffects.svg"), height = 3, width = 8)

# Dimension 6
md6 <- lmer(PC6 ~ Register*Level + (1|Series), data = res.ind, REML = FALSE)
md6Register <- lmer(PC6 ~ Register + (1|Series), data = res.ind, REML = FALSE)
md6Level <- lmer(PC6 ~ Level + (1|Series), data = res.ind, REML = FALSE)

anova(md6, md6Register, md6Level)

tab_model(md6) # Marginal R2 = 0.048

# Plot of fixed effects:
plot_model(md6, 
           type = "est",
           show.intercept = TRUE,
           show.values=TRUE, 
           show.p=TRUE,
           value.offset = .4,
           value.size = 3.5,
           colors = palette[c(1:3,8,7)],
           group.terms = c(1:5,1,1,1,1,2:5,2:5,2:5,2:5), 
           title = "",
           wrap.labels = 40,
           axis.title = "PC6 estimated coefficients") +
  theme_sjplot2() 

#ggsave(here("plots", "TxB_PCA6_lmer_fixedeffects.svg"), height = 3, width = 8)

# Plot of random effects:
plot_model(md6, 
           type = "re", # Option to visualise random effects
           show.values=TRUE, 
           show.p=TRUE,
           value.offset = .4,
           value.size = 3.5,
           color = "bw",
           wrap.labels = 40,
           axis.title = "PC6 estimated coefficients") +
  theme_sjplot2()

#ggsave(here("plots", "TxB_PCA6_lmer_randomeffects.svg"), height = 3, width = 8)

# Boxplot visualisation

res.ind %>% 
  ggplot(aes(x = Register, y = PC2, fill = Level, facet = Register)) +
         geom_boxplot() +
  facet_grid()

ggplot(res.ind, aes(x=Register,y=PC2, fill = Level, colour = Register, facet = Register))+ # Or leave out "colour = Register" to keep the dots in black
  #geom_point(position = position_jitter(width = .15), size = .25)+
# note that here we need to set the x-variable to a numeric variable and bump it to get the boxplots to line up with the rainclouds. 
  geom_boxplot(aes(x = as.numeric(Register), y = PC2), colour = "BLACK") +
  ylab('PC2')+ 
  theme_minimal()+ 
  guides(fill = FALSE, colour = FALSE) +
  scale_colour_manual(values = colours)+
  scale_fill_manual(values = colours)

#ggsave(here("plots", "TxB_PC2_RegisterLevelsBoxplots.svg"), width = 13, height = 8)

```

```{r sessionInfo}
sessionInfo()
```

