---
title: "Reference corpus data preparation for full MDA"
author: "Elen Le Foll"
date: "09/11/2021"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
  editor_options: 
    markdown: 
      wrap: sentence

---

# Set-up

*Built with R `r getRversion()`*  
*Last saved on `r format(Sys.Date(), "%d %B %Y")` at `r format(Sys.time(), "%H:%M")`*

```{r setup, include=TRUE, results = "hide", warning = FALSE, message = FALSE}

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)

knitr::opts_chunk$set(echo = TRUE, tidy = TRUE, message=FALSE, paged.print=TRUE, fig.width = 10, warning=FALSE)

#renv::restore() # Restore the project's dependencies from the lockfile to ensure that same package versions are used as in the original thesis.

library(caret) # For its confusion matrix function
library(dplyr)
library(forcats)
library(here) # For dynamic file paths
library(ggplot2) 
library(PerformanceAnalytics)
library(purrr) # For data wrangling
library(psych) # For various useful stats function
library(suffrager) # For pretty feminist colour palettes :)
library(tidyr)
library(tibble)

```

# MDA of Textbook English vs. Reference corpora

## Importing reference corpora data

### Importing Spoken BNC2014 counts

These counts were computed on the basis of the "John and Jill in Ivybridge" version of the Spoken BNC2014 with added full stops at speaker turns.

```{r SpokenBNC2014}

SpokenBNC2014 <- read.delim(here("MFTE", "Outputs", "SpokenBNC2014_3.1_normed_complex_counts.tsv"), header = TRUE, stringsAsFactors = TRUE)
str(SpokenBNC2014) # Check sanity of data
nrow(SpokenBNC2014) # Should be 1251 files

SpokenBNC2014$Series <- "Spoken BNC2014"
SpokenBNC2014$Level <- "Ref."
SpokenBNC2014$Country <- "Spoken BNC2014"
SpokenBNC2014$Register <- "Spoken BNC2014"

```

### Importing Youth Fiction counts

These counts were computed on the basis of the random samples of approximately 5,000 words of the books of the Youth Fiction corpus.

```{r YouthFiction}

YouthFiction <- read.delim(here("MFTE", "Outputs", "YF_sampled_500_3.1_normed_complex_counts.tsv"), header = TRUE, stringsAsFactors = TRUE)
str(YouthFiction) # Check sanity of data
nrow(YouthFiction) # Should be 1191 files

YouthFiction$Series <- "Youth Fiction"
YouthFiction$Level <- "Ref."
YouthFiction$Country <- "Youth Fiction"
YouthFiction$Register <- "Youth Fiction"

```

### Importing ITTC counts

```{r InfoTeencounts}

InfoTeen <- read.delim(here("MFTE", "Outputs", "InfoTeen_3.1_normed_complex_counts.tsv"), header = TRUE, stringsAsFactors = TRUE)
str(InfoTeen) # Check sanity of data
nrow(InfoTeen) # Should be 1414 files
InfoTeen <- InfoTeen %>% filter(Filename!=".DS_Store" & Filename!="Revision_World_GCSE_10529068_wjec-level-law-past-papers.txt" & Filename!="Revision_World_GCSE_10528474_wjec-level-history-past-papers.txt" & Filename!="Revision_World_GCSE_10528472_edexcel-level-history-past-papers.txt")
# Removes three outlier files which should not have been included in the corpus

InfoTeen$Series <- "Info Teens"
InfoTeen$Level <- "Ref."
InfoTeen$Country <- "Info Teens"
InfoTeen$Register <- "Info Teens"

```


## Merging TEC and reference corpora counts

Due to reasons of space, the results of the five-register dataset were not included in the thesis.
```{r DataMerging-5Reg, eval=FALSE}

TxBcounts <- readRDS(here("FullMDA", "TxBcounts.rds"))

TxBcounts %>% 
  filter(Series=="NGL") %>% 
  group_by(Series, Level) %>% 
  summarise(wordcount = sum(Words))

counts <- bind_rows(TxBcounts, InfoTeen, SpokenBNC2014, YouthFiction, .id = "Corpus") %>% 
  filter(Register != "Poetry")
head(counts); tail(counts)
nrow(counts)

# Convert all character vectors to factors
counts[sapply(counts, is.character)] <- lapply(counts[sapply(counts, is.character)], as.factor)

# Change all NAs to 0
counts[is.na(counts)] <- 0

levels(counts$Corpus)
levels(counts$Corpus) <- list(Textbook.English="1", Informative.Teens="2", Spoken.BNC2014="3", Youth.Fiction="4")
summary(counts$Corpus)
summary(counts$Series)

# Re-order registers
levels(counts$Register)
counts$Register <- factor(counts$Register, levels = c("Conversation", "Fiction", "Informative", "Instructional", "Personal", "Info Teens", "Spoken BNC2014", "Youth Fiction"))

# Wrangle metadata variables
counts$Subcorpus <- counts$Register
levels(counts$Subcorpus) <- c("Textbook Conversation", "Textbook Fiction", "Textbook Informative", "Textbook Instructional", "Textbook Personal", "Info Teens Ref.", "Spoken BNC2014 Ref.", "Youth Fiction Ref.")
summary(counts$Subcorpus)

levels(counts$Register) <- c("Conversation", "Fiction", "Informative", "Instructional", "Personal", "Poetry", "Informative", "Conversation", "Fiction")
summary(counts$Register)

# Re-order variables
colnames(counts)
counts <- counts %>% 
  select(order(names(.))) %>% # Order alphabetically first
  select(Filename, Register, Level, Series, Country, Corpus, Subcorpus, Words, everything()) # Then place the metadata variable at the front of the table

#saveRDS(counts, here("FullMDA", "counts.rds")) # Last saved 9 Feb 2022

```

This is the dataset that is presented in the second half of Chapter 7.
```{r DataMerging-3Reg.only}

TxBcounts <- readRDS(here("FullMDA", "TxBcounts.rds"))

All3Reg <- c("Conversation", "Fiction", "Informative")

TxBcounts3Reg <- TxBcounts %>% 
  filter(Register %in% All3Reg) %>% 
  droplevels(.)

counts <- bind_rows(TxBcounts3Reg, InfoTeen, SpokenBNC2014, YouthFiction, .id = "Corpus")
head(counts); tail(counts)
str(counts)

# Convert all character vectors to factors
counts[sapply(counts, is.character)] <- lapply(counts[sapply(counts, is.character)], as.factor)

# Change all NAs to 0
counts[is.na(counts)] <- 0

levels(counts$Corpus)
levels(counts$Corpus) <- list(Textbook.English="1", Informative.Teens="2", Spoken.BNC2014="3", Youth.Fiction="4")
summary(counts$Corpus)
summary(counts$Series)

# Wrangle metadata variables
counts$Subcorpus <- counts$Register
levels(counts$Subcorpus)
levels(counts$Subcorpus) <- c("Textbook Conversation", "Textbook Fiction", "Info Teens Ref.", "Textbook Informative", "Spoken BNC2014 Ref.", "Youth Fiction Ref.")
summary(counts$Subcorpus)

# Re-order registers
levels(counts$Register)
levels(counts$Register) <- c("Conversation", "Fiction", "Informative", "Informative", "Conversation", "Fiction")
summary(counts$Register)

# Re-order variables
colnames(counts)
counts <- counts %>% 
  select(order(names(.))) %>% # Order alphabetically first
  select(Filename, Register, Level, Series, Country, Corpus, Subcorpus, Words, everything()) # Then place the metadata variable at the front of the table

#saveRDS(counts, here("FullMDA", "counts3Reg.rds")) # Last saved 9 Feb 2022

```

## Data preparation

### Plotting the distributions of all the features
```{r distribution-viz, fig.height=40, warning=FALSE}

ncounts <- readRDS(here("FullMDA", "counts3Reg.rds"))
colnames(ncounts)

# Compare relative frequencies of individual features, e.g., BE as a main verb per FVP (finite verb phrase)
ncounts %>% 
  group_by(Register, Corpus) %>% 
  summarise(median(BEMA), MAD(BEMA))

# Inspired by: https://drsimonj.svbtle.com/quick-plot-of-all-variables

ncounts %>%
  select(-Words) %>% 
  keep(is.numeric) %>% 
  gather() %>% # This function from tidyr converts a selection of variables into two variables: a key and a value. The key contains the names of the original variable and the value the data. This means we can then use the facet_wrap function from ggplot2
  ggplot(aes(value)) +
    theme_bw() +
    facet_wrap(~ key, scales = "free", ncol = 4) +
    scale_x_continuous(expand=c(0,0)) +
    scale_y_continuous(limits = c(0,NA)) +
    geom_histogram(aes(y = ..density..), bins = 30, colour= "black", fill = "grey") +
    geom_density(colour = "darkred", weight = 2, fill="darkred", alpha = .4)

#ggsave(here("Plots", "DensityPlotsAllVariables.svg"), width = 15, height = 49)

ncounts %>%
  select(-Words) %>% 
  keep(is.numeric) %>% 
  gather() %>% # This function from tidyr converts a selection of variables into two variables: a key and a value. The key contains the names of the original variable and the value the data. This means we can then use the facet_wrap function from ggplot2
  ggplot(aes(value)) +
    theme_bw() +
    facet_wrap(~ key, scales = "free", ncol = 4) +
    scale_x_continuous(expand=c(0,0)) +
    geom_histogram(bins = 30, colour= "darkred", fill = "darkred", alpha = 0.5)

#ggsave(here("Plots", "HistogramPlotsAllVariables.svg"), width = 20, height = 45)

```

### Feature removal due to low text frequency
```{r feature-removal-low-frequency}

# For MDA with five TEC registers
#ncounts <- readRDS(here("FullMDA", "counts.rds"))

# For MDA with three TEC registers
ncounts <- readRDS(here("FullMDA", "counts3Reg.rds"))

colnames(ncounts)

# Removal of meaningless feature: CD because numbers as digits were mostly removed from the textbooks, LIKE and SO because they are dustbin categories
ncounts <- ncounts %>% 
  select(-c(CD, LIKE, SO))

# Combine low frequency features into meaningful groups whenever this makes linguistic sense
ncounts <- ncounts %>% 
  mutate(JJPR = JJPR + ABLE, ABLE = NULL) %>% 
  mutate(PASS = PGET + PASS, PGET = NULL) %>% 
  mutate(TPP3 = TPP3S + TPP3P, TPP3P = NULL, TPP3S = NULL) %>% 
  mutate(FQTI = FREQ + TIME, FREQ = NULL, TIME = NULL)

zero_features <- as.data.frame(round(colSums(ncounts==0)/nrow(ncounts)*100, 2)) # Percentage of texts with 0 occurrences of each feature
colnames(zero_features) <- "Percentage_with_zero"
zero_features %>% 
  filter(!is.na(zero_features)) %>% 
  rownames_to_column() %>% 
  arrange(Percentage_with_zero) %>% 
  filter(Percentage_with_zero > 66.6)

zero_features <- as.data.frame(round(colSums(ncounts>0)/nrow(ncounts)*100, 2)) # Percentage of texts >0 occurrences of each feature
colnames(zero_features) <- "Percentage_above_zero"
zero_features %>% rownames_to_column() %>% filter(!is.na(zero_features)) %>% arrange(desc(Percentage_above_zero)) 

docfreq.too.low <- zero_features %>% filter(!is.na(zero_features)) %>% subset(Percentage_above_zero < 33.3) %>% rownames_to_column() %>% select(rowname) # Select all variables with a document frequency of at least 40%.
docfreq.too.low

ncounts <- select(ncounts, -one_of(docfreq.too.low$rowname)) # Drop these variables
colnames(ncounts)
ncol(ncounts)-8 # Number of linguistic features remaining

# With five TEC registers
#saveRDS(ncounts, here("FullMDA", "ncounts2.rds")) # Last saved 18 November 2021

# With three TEC registers
#saveRDS(ncounts, here("FullMDA", "ncounts2_3Reg.rds")) # Last saved 9 Feb 2022

```

### Standardising normalised counts and identifying potential outliers

"As an alternative to removing very sparse feature, we apply a signed logarithmic transformation to deskew the feature distributions." (Neumann & Evert)

```{r standardisation-outliers}

# First scale the normalised counts (z-standardisation) to be able to compare the various features
ncounts %>%
  select(-Words) %>% 
  keep(is.numeric) %>% 
  scale() ->
  zcounts

boxplot(zcounts, las = 3, main = "z-scores") # Slow

# If necessary, remove any outliers at this stage.
colnames(ncounts)
data <- cbind(ncounts[,1:8], as.data.frame(zcounts))
str(data)
nrow(data)

outliers <- data %>% 
 filter(if_any(where(is.numeric) & !Words,  .fns = function(x){x > 8}))  %>% 
  select(Filename, Corpus, Series, Register, Level, Words)
outliers

outliers %>% select(Filename)

# Checking that outlier texts are not particularly long or short texts
summary(outliers$Words)
histogram(outliers$Words, breaks = 30)
summary(data$Words)
# Distribution of outlier texts
summary(outliers$Corpus)

# Manually checking a sample of these outliers:

# Encyclopedia_Kinds_au_10085347_Nobel_Prize_in_Chemistry.txt is essentially a list of Nobel prize winners but with some additional information. Hence a good representative of the type of texts of the ITTC.
# Solutions_Elementary_ELF_Spoken_0013 --> Has a lot of "going to" constructions because they are learnt in this chapter but is otherwise a well-formed text.
# Teen_Kids_News_10403972_a-brief-history-of-white-house-weddings --> No issues
# Teen_Kids_News_10403301_golden-globe-winners-2019-the-complete-list --> Similar to the Nobel prize laureates text.
# Revision_World_GCSE_10528123_gender-written-textual-analysis-framework --> Text includes bullet points tokenised as the letter "o" but otherwise a fairly typical informative text.

# Removing the outliers
ncounts <- ncounts %>% 
  filter(!Filename %in% outliers$Filename)

nrow(ncounts)

#saveRDS(ncounts, here("FullMDA", "ncounts3_3Reg.rds")) # Last saved 9 Feb 2022

ncounts %>%
  select(-Words) %>% 
  keep(is.numeric) %>% 
  scale() ->
  zcounts

nrow(zcounts)
boxplot(zcounts, las = 3, main = "z-scores") # Slow to open!

```

### Transforming the features to (partially) deskew these distributions

```{r signed.log.transformation}

signed.log <- function(x) {sign(x)*log(abs(x)+1)}

zlogcounts <- signed.log(zcounts) # Standardise first, then signed log transform

boxplot(zlogcounts, las=3, main="log-transformed z-scores")

# With three TEC registers
#saveRDS(zlogcounts, here("FullMDA", "zlogcounts_3Reg.rds")) # Last saved 9 Feb 2022

# With five TEC registers
#saveRDS(zlogcounts, here("FullMDA", "zlogcounts.rds")) # Last saved 18 November

zlogcounts %>%
  as.data.frame() %>% 
  gather() %>% # This function from tidyr converts a selection of variables into two variables: a key and a value. The key contains the names of the original variable and the value the data. This means we can then use the facet_wrap function from ggplot2
  ggplot(aes(value)) +
  theme_bw() +
  facet_wrap(~ key, scales = "free", ncol = 4) +
  scale_x_continuous(expand=c(0,0)) +
  scale_y_continuous(limits = c(0,NA)) +
  geom_histogram(aes(y = ..density..), bins = 30, colour= "black", fill = "grey") +
  geom_density(colour = "darkred", weight = 2, fill="darkred", alpha = .4)

#ggsave(here("Plots", "DensityPlotsAllVariablesSignedLog.svg"), width = 15, height = 49)

```


### Merging of data for MDA

```{r import-final-ncounts-data}

# With five TEC registers 
# zlogcounts <- readRDS(here("FullMDA", "zlogcounts.rds")) 
# nrow(zlogcounts)
# colnames(zlogcounts)
# 
# ncounts <- readRDS(here("FullMDA", "ncounts2.rds"))
# nrow(ncounts)
# colnames(ncounts)
# 
# data <- cbind(ncounts[,1:7], as.data.frame(zlogcounts))
# str(data)
# 
#saveRDS(data, here("FullMDA", "datazlogcounts.rds")) # Last saved 18 November

# With three TEC registers
zlogcounts <- readRDS(here("FullMDA", "zlogcounts_3Reg.rds")) 
nrow(zlogcounts)
colnames(zlogcounts)

ncounts <- readRDS(here("FullMDA", "ncounts3_3Reg.rds"))
nrow(ncounts)
colnames(ncounts)

data <- cbind(ncounts[,1:8], as.data.frame(zlogcounts))
colnames(data)

#saveRDS(data, here("FullMDA", "datazlogcounts_3Reg.rds")) # Last saved 9 Feb 2022

```

## Performing the PCA of Textbook English vs. Reference corpora
### Quick import
```{r prepare-data}

# With five TEC registers
# data <- readRDS(here("FullMDA", "datazlogcounts.rds"))

# With three TEC registers
data <- readRDS(here("FullMDA", "datazlogcounts_3Reg.rds"))

summary(data$Corpus)
summary(data$Subcorpus)

# This rearranges the levels in the desired order for the plot legends:
data <- data %>% 
  mutate(Subcorpus = fct_relevel(Subcorpus, "Info Teens Ref.", after = 9))

```

## Testing factorability of data

### Correlation matrix: only for data exploration, not used to exclude variables
```{r corrsimple-function, eval=FALSE}

# From: https://towardsdatascience.com/how-to-create-a-correlation-matrix-with-too-many-variables-309cc0c0a57
## Function adapated to my needs ##

colnames(data)

corr <- cor(data[9:ncol(data)])
#prepare to drop duplicates and correlations of 1     
corr[lower.tri(corr,diag=TRUE)] <- NA 
#drop perfect correlations
corr[corr == 1] <- NA   
#turn into a 3-column table
corr <- as.data.frame(as.table(corr))
#remove the NA values from above 
corr <- na.omit(corr)   

#Uninteresting variable correlations?
lowcor <- subset(corr, abs(Freq) < 0.3)
#lowcor %>% filter(Var2=="CC"|Var1=="CC") %>% round(Freq, 2)
#select significant correlations 
corr <- subset(corr, abs(Freq) > 0.3)
#sort by highest correlation
corr <- corr[order(-abs(corr$Freq)),]   
#see which variables might be eliminated: the ones with correlation > 0.3
eliminate <- as.data.frame((summary(corr$Var1) + summary(corr$Var2)))
(LowcCommunality <- eliminate %>% filter(`(summary(corr$Var1) + summary(corr$Var2))`==0))

# Potentially problematic collinear variables that may need to be removed:
highcor <- subset(corr, abs(Freq) > 0.95)
highcor

#variables which are retained
corr$Var1 <- droplevels(corr$Var1)
corr$Var2 <- droplevels(corr$Var2)
features <- unique(c(levels(corr$Var1), levels(corr$Var2))) 
features # 68 variables
#turn corr back into matrix in order to plot with corrplot
mtx_corr <- reshape2::acast(corr, Var1~Var2, value.var="Freq")
  
#plot correlations in a manageable way
library(corrplot)
plot.margin=unit(c(0,0,0,0), "mm")
corrplot(mtx_corr, is.corr=FALSE, tl.col="black", na.label=" ", tl.cex = 0.5)
#save as SVG with Rstudio e.g. 1000 x 1000
```

### Visualisation of feature correlations
```{r heatmap}

# Simple heatmap in base R (inspired by Stephanie Evert's SIGIL code)
cor.colours <- c(
  hsv(h=2/3, v=1, s=(10:1)/10), # blue = negative correlation 
  rgb(1,1,1), # white = no correlation 
  hsv(h=0, v=1, s=(1:10/10))) # red = positive correlation

#png(here("Plots", "heatmapzlogcounts.png"), width = 30, height= 30, units = "cm", res = 300)
heatmap(cor(zlogcounts), 
        symm=TRUE, 
        zlim=c(-1,1), 
        col=cor.colours, 
        margins=c(7,7))
dev.off()

```

### MSA, communalities and scree plot
```{r factorability-screeplot}

# Eliminate highly collinear variable
cor(data$VPRT, data$VBD)

data <- data %>% 
  select(-c(VPRT))

colnames(data)
kmo <- KMO(data[,9:ncol(data)])
kmo # # Overall MSA = 0.95 
kmo$MSAi[order(kmo$MSAi)] # All features have individual MSAs of > 0.5 (but only because TPP3P was merged with TPP3S earlier on)

#png(here("Plots", "screeplot-TEC-Ref_3Reg.png"), width = 20, height= 12, units = "cm", res = 300)
scree(data[,9:ncol(data)], factors = FALSE, pc = TRUE) # 6 components
dev.off()

# Perform PCA
pca1 <- psych::principal(data[9:ncol(data)], 
                         nfactors = 6)
pca1$loadings

pca1$communality %>% sort(.) # If features with communalities of < 0.2 are removed, we remove TIME (therefore merged TIME and FREQ further up the line)

# Final number of features
ncol(data)-6
# Final number of texts
nrow(data)

#saveRDS(data, here("FullMDA", "dataforPCA.rds")) # Last saved on 9 Feb 2022

```

```{r sessionInfo}
sessionInfo()
```
