---
title: "Tagger evaluation"
author: "Elen Le Foll"
date: "01/02/2022"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
    #code_folding: hide # enables you to include R code but have it hidden by default. Users can then choose to show hidden R code chunks either individually or document wide.
    #keep_md: true
    md_extensions: +bracketed_spans      
      
bibliography: "packages.bib"
nocite: '@*'
      
---

This script is part of the Online Appendix to my PhD thesis.

Please cite as: 
Le Foll, Elen. 2022. Textbook English: A Corpus-Based Analysis of the Language of EFL textbooks used in Secondary Schools in France, Germany and Spain. PhD thesis. Osnabrück University.

For more information, see: https://elenlefoll.github.io/TextbookEnglish/

Please note that the plot dimensions in this notebook have been optimised for the print version of the thesis.


### Set-up

*Built with R `r getRversion()`*  

# Tagger evaluation on TEC data

## Data import from evaluation files

These chunks import the data directly from the Excel files in which the manual tag check and corrections was performed.

```{r setup, include=TRUE, results = "hide", warning = FALSE, message = FALSE}

knitr::opts_chunk$set(echo = TRUE, tidy = TRUE, message=FALSE, paged.print=TRUE, fig.width = 10, warning=FALSE)

#renv::restore() # Restore the project's dependencies from the lockfile to ensure that same package versions are used as in the original thesis.

library(caret) # For computing confusion matrices
library(harrypotter) # Only for colour scheme
library(here) # For path management
library(paletteer) # For nice colours
library(readxl) # For the direct import of Excel files
library(tidyverse) # For everything else!

```

```{r TxBcounts}

# Function to import and wrangle the evaluation data from the Excel files in which the manual evaluation was conducted
importEval3 <- function(file, fileID, register, corpus) {
  Tag1 <- file %>% 
  add_column(FileID = fileID, Register = register, Corpus = corpus) %>%
  select(FileID, Corpus, Register, Output, Tokens, Tag1, Tag1Gold) %>% 
  rename(Tag = Tag1, TagGold = Tag1Gold, Token = Tokens) %>% 
  mutate(Evaluation = ifelse(is.na(TagGold), TRUE, FALSE)) %>% 
  mutate(TagGold = ifelse(is.na(TagGold), as.character(Tag), as.character(TagGold))) %>%
  filter(!is.na(Tag)) %>% 
  mutate_if(is.character, as.factor)
  
  Tag2 <- file %>% 
  add_column(FileID = fileID, Register = register, Corpus = corpus) %>%
  select(FileID, Corpus, Register, Output, Tokens, Tag2, Tag2Gold) %>% 
  rename(Tag = Tag2, TagGold = Tag2Gold, Token = Tokens) %>% 
  mutate(Evaluation = ifelse(is.na(TagGold), TRUE, FALSE)) %>% 
  mutate(TagGold = ifelse(is.na(TagGold), as.character(Tag), as.character(TagGold))) %>%
  filter(!is.na(Tag)) %>% 
  mutate_if(is.character, as.factor)

Tag3 <- file %>% 
  add_column(FileID = fileID, Register = register, Corpus = corpus) %>%
  select(FileID, Corpus, Register, Output, Tokens, Tag3, Tag3Gold) %>% 
  rename(Tag = Tag3, TagGold = Tag3Gold, Token = Tokens) %>% 
  mutate(Evaluation = ifelse(is.na(TagGold), TRUE, FALSE)) %>% 
  mutate(TagGold = ifelse(is.na(TagGold), as.character(Tag), as.character(TagGold))) %>%
  filter(!is.na(Tag)) %>% 
  mutate_if(is.character, as.factor)

output <- rbind(Tag1, Tag2, Tag3) %>% 
  mutate(across(where(is.factor), str_remove_all, pattern = fixed(" "))) %>% # Removes all white spaces which are found in the excel files
  filter(!is.na(Output)) %>% 
  mutate_if(is.character, as.factor)
}

# Second function to import and wrangle the evaluation data for Excel files with four tag columns as opposed to three
importEval4 <- function(file, fileID, register, corpus) {
  Tag1 <- file %>% 
  add_column(FileID = fileID, Register = register, Corpus = corpus) %>%
  select(FileID, Corpus, Register, Output, Tokens, Tag1, Tag1Gold) %>% 
  rename(Tag = Tag1, TagGold = Tag1Gold, Token = Tokens) %>% 
  mutate(Evaluation = ifelse(is.na(TagGold), TRUE, FALSE)) %>% 
  mutate(TagGold = ifelse(is.na(TagGold), as.character(Tag), as.character(TagGold))) %>%
  filter(!is.na(Tag)) %>% 
  mutate_if(is.character, as.factor)
  
  Tag2 <- file %>% 
  add_column(FileID = fileID, Register = register, Corpus = corpus) %>%
  select(FileID, Corpus, Register, Output, Tokens, Tag2, Tag2Gold) %>% 
  rename(Tag = Tag2, TagGold = Tag2Gold, Token = Tokens) %>% 
  mutate(Evaluation = ifelse(is.na(TagGold), TRUE, FALSE)) %>% 
  mutate(TagGold = ifelse(is.na(TagGold), as.character(Tag), as.character(TagGold))) %>%
  filter(!is.na(Tag)) %>% 
  mutate_if(is.character, as.factor)

Tag3 <- file %>% 
  add_column(FileID = fileID, Register = register, Corpus = corpus) %>%
  select(FileID, Corpus, Register, Output, Tokens, Tag3, Tag3Gold) %>% 
  rename(Tag = Tag3, TagGold = Tag3Gold, Token = Tokens) %>% 
  mutate(Evaluation = ifelse(is.na(TagGold), TRUE, FALSE)) %>% 
  mutate(TagGold = ifelse(is.na(TagGold), as.character(Tag), as.character(TagGold))) %>%
  filter(!is.na(Tag)) %>% 
  mutate_if(is.character, as.factor)

Tag4 <- file %>% 
  add_column(FileID = fileID, Register = register, Corpus = corpus) %>%
  select(FileID, Corpus, Register, Output, Tokens, Tag4, Tag4Gold) %>% 
  rename(Tag = Tag4, TagGold = Tag4Gold, Token = Tokens) %>% 
  mutate(Evaluation = ifelse(is.na(TagGold), TRUE, FALSE)) %>% 
  mutate(TagGold = ifelse(is.na(TagGold), as.character(Tag), as.character(TagGold))) %>%
  filter(!is.na(Tag)) %>% 
  mutate_if(is.character, as.factor)

output <- rbind(Tag1, Tag2, Tag3, Tag4) %>% 
  mutate(across(where(is.factor), str_remove_all, pattern = fixed(" "))) %>% # Removes all white spaces which are found in the excel files
  filter(!is.na(Tag)) %>% 
  mutate_if(is.character, as.factor)

}

# Function to decide which of the two above functions should be used
importEval <- function(file, fileID, register, corpus) { 
  if(sum(!is.na(file$Tag4)) > 0) {
    output = importEval4(file = file, fileID = fileID, register = register, corpus = corpus)
  }
  else{
    output = importEval3(file = file, fileID = fileID, register = register, corpus = corpus)
  }
}

Solutions_Intermediate_Spoken_0032 <- importEval(file = read_excel(here("MFTE", "Evaluation", "Solutions_Intermediate_Spoken_0032_Evaluation.xlsx")), fileID = "Solutions_Intermediate_Spoken_0032", register = "Conversation", corpus = "TEC-Sp")

HT_5_Poetry_0001 <- importEval(file = read_excel(here("MFTE", "Evaluation", "HT_5_Poetry_0001_Evaluation.xlsx")), fileID = "HT_5_Poetry_0001", register = "Poetry", corpus = "TEC-Fr")

Achievers_A1_Informative_0006 <- importEval(file = read_excel(here("MFTE", "Evaluation", "Achievers_A1_Informative_0006_Evaluation.xlsx")), fileID = "Achievers_A1_Informative_0006", register = "Informative", corpus = "TEC-Sp")

New_GreenLine_5_Personal_0003 <- importEval(file = read_excel(here("MFTE", "Evaluation", "New_GreenLine_5_Personal_0003_Evaluation.xlsx")), fileID = "New_GreenLine_5_Personal_0003", register = "Personal communication", corpus = "TEC-Ger")

Piece_of_cake_3e_Instructional_0006 <- importEval(file = read_excel(here("MFTE", "Evaluation", "Piece_of_cake_3e_Instructional_0006_Evaluation.xlsx")), fileID = "Piece_of_cake_3e_Instructional_0006", register = "Instructional", corpus = "TEC-Fr")

Access_4_Narrative_0006 <- importEval(file = read_excel(here("MFTE", "Evaluation", "Access_4_Narrative_0006_Evaluation.xlsx")), fileID = "Access_4_Narrative_0006", register = "Fiction", corpus = "TEC-Ger")

BNCBFict_b2 <- importEval(file = read_excel(here("MFTE", "Evaluation", "BNCBFict_b2.xlsx")), fileID = "BNCBFict_b2", register = "fiction", corpus = "BNC2014")

BNCBFict_m54 <- importEval(file = read_excel(here("MFTE", "Evaluation", "BNCBFict_m54.xlsx")), fileID = "BNCBFict_m54", register = "fiction", corpus = "BNC2014")

BNCBFict_e27 <- importEval(file = read_excel(here("MFTE", "Evaluation", "BNCBFict_e27.xlsx")), fileID = "BNCBFict_e27", register = "fiction", corpus = "BNC2014")

BNCBMass16 <- importEval(file = read_excel(here("MFTE", "Evaluation", "BNCBMass16.xlsx")), fileID = "BNCBMass16", register = "news", corpus = "BNC2014")

BNCBMass23 <- importEval(file = read_excel(here("MFTE", "Evaluation", "BNCBMass23.xlsx")), fileID = "BNCBMass23", register = "news", corpus = "BNC2014")

BNCBReg111 <- importEval(file = read_excel(here("MFTE", "Evaluation", "BNCBReg111.xlsx")), fileID = "BNCBReg111", register = "news", corpus = "BNC2014")

BNCBReg750 <- importEval(file = read_excel(here("MFTE", "Evaluation", "BNCBReg750.xlsx")), fileID = "BNCBReg750", register = "news", corpus = "BNC2014")

BNCBSer486 <- importEval(file = read_excel(here("MFTE", "Evaluation", "BNCBSer486.xlsx")), fileID = "BNCBSer486", register = "news", corpus = "BNC2014")

BNCBSer562 <- importEval(file = read_excel(here("MFTE", "Evaluation", "BNCBSer562.xlsx")), fileID = "BNCBSer562", register = "news", corpus = "BNC2014")

BNCBEBl8 <- importEval(file = read_excel(here("MFTE", "Evaluation", "BNCBEBl8.xlsx")), fileID = "BNCBEBl8", register = "internet", corpus = "BNC2014")

BNCBEFor32 <- importEval(file = read_excel(here("MFTE", "Evaluation", "BNCBEFor32.xlsx")), fileID = "BNCBEFor32", register = "internet", corpus = "BNC2014")

S2DD <- importEval(file = read_excel(here("MFTE", "Evaluation", "S2DD.xlsx")), fileID = "S2DD", register = "spoken", corpus = "BNC2014")

S3AV <- importEval(file = read_excel(here("MFTE", "Evaluation", "S3AV.xlsx")), fileID = "S3AV", register = "spoken", corpus = "BNC2014")

SEL5 <- importEval(file = read_excel(here("MFTE", "Evaluation", "SEL5.xlsx")), fileID = "SEL5", register = "spoken", corpus = "BNC2014")

SVLK <- importEval(file = read_excel(here("MFTE", "Evaluation", "SVLK.xlsx")), fileID = "SVLK", register = "spoken", corpus = "BNC2014")

SZXQ <- importEval(file = read_excel(here("MFTE", "Evaluation", "SZXQ.xlsx")), fileID = "SZXQ", register = "spoken", corpus = "BNC2014")

TaggerEval <- rbind(Solutions_Intermediate_Spoken_0032, HT_5_Poetry_0001, Achievers_A1_Informative_0006, New_GreenLine_5_Personal_0003, Piece_of_cake_3e_Instructional_0006, Access_4_Narrative_0006, BNCBEBl8, BNCBFict_b2, BNCBFict_m54, BNCBFict_e27, BNCBEFor32, BNCBMass16, BNCBMass23, BNCBReg111, BNCBReg750, BNCBSer486, BNCBSer562, S2DD, S3AV, SEL5, SVLK, SZXQ) 

summary(TaggerEval)

```

## Implement changes made to MFTE since evaluation

```{r feature merging}

TaggerEval <- TaggerEval %>% 
  mutate(Tag = ifelse(Tag == "PHC", "CC", as.character(Tag))) %>% 
  mutate(TagGold = ifelse(TagGold == "PHC", "CC", as.character(TagGold))) %>% 
  mutate(Tag = ifelse(Tag == "QLIKE", "LIKE", as.character(Tag))) %>% 
  mutate(TagGold = ifelse(TagGold == "QLIKE", "LIKE", as.character(TagGold))) %>% 
  mutate(Tag = ifelse(Tag == "TO", "IN", as.character(Tag))) %>% 
  mutate(TagGold = ifelse(TagGold == "TO", "IN", as.character(TagGold))) %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(Evaluation = ifelse(as.character(Tag) == as.character(TagGold), TRUE, FALSE))

head(TaggerEval) # Check sanity of data
summary(TaggerEval) # Check sanity of data

#saveRDS(TaggerEval, here("MFTE", "Evaluation", "MFTE_PhD_Evaluation_Results.rds")) # Last saved 10 Nov 2021

#write.csv(TaggerEval, here("MFTE", "Evaluation", "MFTE_PhD_Evaluation_Results.csv")) # Last saved 10 Nov 2021

```

### Quick data import

```{r quick-import}

TaggerEval <- readRDS(here("MFTE", "Evaluation", "MFTE_PhD_Evaluation_Results.rds")) 
summary(TaggerEval)

```

## Estimating MFTE accuracy for TEC only

In this chunk, I calculate the recall and precision rates of each feature, ignoring unclear tokens and all punctuation and symbols.

```{r tagger-accuracy-TEC}

# Total number of TEC tags manually checked
TaggerEval %>% filter(Corpus %in% c("TEC-Fr", "TEC-Ger", "TEC-Sp")) %>% nrow()

# Number of UNCLEAR evaluation tags
TaggerEval %>% 
  filter(Corpus %in% c("TEC-Fr", "TEC-Ger", "TEC-Sp")) %>% 
  filter(TagGold == "UNCLEAR") %>% 
  nrow() # 0 in TEC sample

data <- TaggerEval %>% 
  filter(Corpus %in% c("TEC-Fr", "TEC-Ger", "TEC-Sp")) %>% 
  filter(TagGold != "UNCLEAR") %>% 
  filter(Tag %in% c(str_extract(Tag, "[A-Z0-9]+"))) %>% # Remove punctuation tags which are uninteresting here.
  filter(Tag != "SYM" & Tag != "``") %>% 
  droplevels(.) %>% 
  mutate(Tag = factor(Tag, levels = union(levels(Tag), levels(TagGold)))) %>% # Ensure that the factor levels are the same for the next caret operation
  mutate(TagGold = factor(TagGold, levels = union(levels(Tag), levels(TagGold))))

# Spot gold tag corrections that are not actually errors (should return zero rows if all is well)
data[data$Tag==data$TagGold & data$Evaluation == FALSE,] %>% as.data.frame()

# Total number of false tags
summary(data$Evaluation)

cm <- caret::confusionMatrix(data$Tag, data$TagGold) # Create confusion matrix
cm$overall # Note that is not very representative because it includes tags which are not intended for use in the MDA studies, e.g., LS and FW.

# Accuracy metrics per feature: recall, precision and f1
cm$byClass[,5:7]

```

## MFTE accuracy for reference corpora or comparable

### Conversation

```{r tagger-accuracy-SpokenBNC2014}

# Number of UNCLEAR evaluation tags
TaggerEval %>% 
  filter(Register == "spoken") %>% 
  filter(TagGold == "UNCLEAR") %>% 
  nrow() # 7 in Spoken BNC2014 sample

data <- TaggerEval %>% 
  filter(Register == "spoken") %>% 
  filter(TagGold != "UNCLEAR") %>% 
  filter(Tag %in% c(str_extract(Tag, "[A-Z0-9]+"))) %>% # Remove all punctuation tags which are uninteresting here.
  droplevels(.) %>% 
  mutate(Tag = factor(Tag, levels = union(levels(Tag), levels(TagGold)))) %>% # Ensure that the factor levels are the same for the next caret operation
  mutate(TagGold = factor(TagGold, levels = union(levels(Tag), levels(TagGold))))

# Spot gold tag corrections that are not actually errors (should return zero rows if all is well)
data[data$Tag==data$TagGold & data$Evaluation == FALSE,] %>% as.data.frame()

# Total number of false tags
summary(data$Evaluation)

cm <- caret::confusionMatrix(data$Tag, data$TagGold) # Create confusion matrix
cm$overall 

# Accuracy metrics per feature: recall, precision and f1
cm$byClass[,5:7]

```

### Fiction

```{r tagger-accuracy-Fiction}

# Number of UNCLEAR evaluation tags
TaggerEval %>% 
  filter(Register == "fiction") %>% 
  filter(TagGold == "UNCLEAR") %>% 
  nrow() # 0 in BNC2014 Baby+ fiction subsample

data <- TaggerEval %>% 
  filter(Register == "fiction") %>% 
  filter(TagGold != "UNCLEAR") %>% 
  filter(Tag %in% c(str_extract(Tag, "[A-Z0-9]+"))) %>% # Remove all punctuation tags which are uninteresting here.
  filter(Tag != "SYM" & Tag != "``") %>% 
  droplevels(.) %>% 
  mutate(Tag = factor(Tag, levels = union(levels(Tag), levels(TagGold)))) %>% # Ensure that the factor levels are the same for the next caret operation
  mutate(TagGold = factor(TagGold, levels = union(levels(Tag), levels(TagGold))))

# Spot gold tag corrections that are not actually errors (should return zero rows if all is well)
data[data$Tag==data$TagGold & data$Evaluation == FALSE,] %>% as.data.frame()

# Total number of false tags
summary(data$Evaluation)

cm <- caret::confusionMatrix(data$Tag, data$TagGold) # Create confusion matrix
cm$overall 

# Accuracy metrics per feature: recall, precision and f1
cm$byClass[,5:7]

```

### Informative

```{r tagger-accuracy-Informative}

# Number of files and tags included in this part of the evaluation (intended to match the ITTC data)
TaggerEval %>% 
  filter(Register == "news" | FileID %in% c("BNCBEFor32", "BNCBEBl8")) %>% 
  group_by(FileID) %>% 
  count() 

# Number of UNCLEAR evaluation tags
TaggerEval %>% 
  filter(Register == "news" | FileID %in% c("BNCBEFor32", "BNCBEBl8")) %>% 
  filter(TagGold == "UNCLEAR") %>% 
  nrow() # 8

data <- TaggerEval %>% 
  filter(Register == "news" | FileID %in% c("BNCBEFor32", "BNCBEBl8")) %>% 
  filter(TagGold != "UNCLEAR") %>% 
  filter(Tag %in% c(str_extract(Tag, "[A-Z0-9]+"))) %>% # Remove all punctuation tags which are uninteresting here.
  filter(Tag != "SYM" & Tag != "``") %>% 
  droplevels(.) %>% 
  mutate(Tag = factor(Tag, levels = union(levels(Tag), levels(TagGold)))) %>% # Ensure that the factor levels are the same for the next caret operation
  mutate(TagGold = factor(TagGold, levels = union(levels(Tag), levels(TagGold))))

# Spot gold tag corrections that are not actually errors (should return zero rows if all is well)
data[data$Tag==data$TagGold & data$Evaluation == FALSE,] %>% as.data.frame()

# Total number of false tags
summary(data$Evaluation)

cm <- caret::confusionMatrix(data$Tag, data$TagGold) # Create confusion matrix
cm$overall 

# Accuracy metrics per feature: recall, precision and f1
cm$byClass[,5:7]

```

## Estimating overall MFTE accuracy for corpora used in thesis

```{r overall-accuracy}

# Number of tags evaluated per file
TaggerEval %>% group_by(FileID) %>% count(.) %>% arrange(desc(n)) %>% as.data.frame()

# Number of UNCLEAR tokens
TaggerEval %>% filter(TagGold == "UNCLEAR") %>% nrow()

# Tagger evaluation
summary(TaggerEval$Evaluation)

data <- TaggerEval %>% 
  filter(TagGold != "UNCLEAR") %>% 
  filter(Tag %in% c(str_extract(Tag, "[A-Z0-9]+"))) %>% # Remove all punctuation tags which are uninteresting here.
  filter(Tag != "SYM" & Tag != "``") %>% 
  filter(TagGold != "SYM" & TagGold != "``") %>% 
  droplevels(.) %>% 
  mutate(Tag = factor(Tag, levels = union(levels(Tag), levels(TagGold)))) %>% # Ensure that the factor levels are the same for the next caret operation
  mutate(TagGold = factor(TagGold, levels = union(levels(Tag), levels(TagGold))))

cm <- caret::confusionMatrix(data$Tag, data$TagGold) # Create confusion matrix
cm$overall 

# Quick summary of results: recall, precision and f1
cm$byClass[,5:7]

# Generate a better formatted results table for export: recall, precision and f1
confusion_matrix <- cm$table
total <- sum(confusion_matrix)
number_of_classes <- nrow(confusion_matrix)
correct <- diag(confusion_matrix)
# sum all columns
total_actual_class <- apply(confusion_matrix, 2, sum)
# sum all rows
total_pred_class <- apply(confusion_matrix, 1, sum)
# Precision = TP / all that were predicted as positive
precision <- correct / total_pred_class
# Recall = TP / all that were actually positive
recall <- correct / total_actual_class
# F1
f1 <- (2 * precision * recall) / (precision + recall)
# create data frame to output results
results <- data.frame(precision, recall, f1, total_actual_class)
results

```

````{r plot-accuracy, fig.height = 20}

resultslong <- results %>% 
  drop_na() %>% 
  mutate(tag = row.names(.)) %>% 
  filter(tag != "NULL" & tag != "SYM" & tag != "OCR" & tag != "FW" & tag != "USEDTO") %>% 
  rename(n = total_actual_class) %>% 
  pivot_longer(cols = c("precision", "recall", "f1"), names_to = "metric", values_to = "value") %>% 
  mutate(metric = factor(metric, levels = c("precision", "recall", "f1")))

summary(resultslong$n)

ggplot(resultslong, aes(y = reorder(tag, desc(tag)), x = value, group = metric, colour = n)) +
  geom_point(size = 2) +
  ylab("") +
  xlab("") +
  facet_wrap(~ metric) +
  scale_color_paletteer_c("harrypotter::harrypotter", trans = "log", breaks = c(1,10, 100, 1000), labels = c(1,10, 100, 1000), name = "# tokens \nmanually\nevaluated") +
  theme_bw() +
  theme(panel.grid.major.y = element_line(colour = "darkgrey")) +
  theme(legend.position = "right")

#ggsave(here("Plots", "TaggerAccuracyPlot.svg"), width = 7, height = 12)

```

## Exploring tagger errors

```{r errors}

# Adding an error tag with the incorrectly assigned tag and underscore and then the correct "gold" label
errors <- TaggerEval %>% 
  filter(Evaluation=="FALSE") %>% 
  filter(TagGold != "UNCLEAR") %>% 
  mutate(Error = paste(Tag, TagGold, sep = " -> ")) 

# Total number of errors
nrow(errors) # 817

FreqErrors <- errors %>% 
  #filter(Corpus %in% c("TEC-Fr", "TEC-Ger", "TEC-Sp")) %>% 
  count(Error) %>% 
  arrange(desc(n))

# Number of error types that only occur once
FreqErrors %>% 
  filter(n == 1) %>% 
  nrow()

# Total number of error types
nrow(FreqErrors)

FreqErrors %>% 
  #filter(n > 10) %>% 
  print.data.frame() 

errors %>% 
  filter(Error == "THSC -> THRC") %>% 
  select(FileID, Output, Tag, TagGold) %>% 
  print(n=30)

errors %>% 
  filter(Error == "NN -> JJAT") %>% 
  select(-Output, -Corpus, -Tag, -TagGold) %>% 
  filter(grepl(x = Token, pattern = "[A-Z]+.")) %>% 
  print.data.frame() 

errors %>% 
  filter(Error %in% c("NN -> VB", "VB -> NN", "NN -> VPRT", "VPRT -> NN")) %>% 
  count(Token) %>% 
  arrange(desc(n)) %>% 
  print.data.frame() 

errors %>% 
  filter(Error == "ACT -> NULL") %>% 
  count(Token) %>% 
  arrange(desc(n)) %>% 
  print.data.frame() 

```

# Package used in this script
```{r package-citations}

#packages.bib <- sapply(1:length(loadedNamespaces()), function(i) toBibtex(citation(loadedNamespaces()[i])))

knitr::write_bib(c(.packages(), "knitr"), "packages.bib")

sessionInfo()

```


