---
title: "Progressives Morphosyntactic Analysis of Textbook Conversation"
author: "Elen Le Foll"
date: "09/05/2019"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false

---

## Set-up

```{r setup, include=TRUE, results = "hide", warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(here)
library(ggplot2)
library(gridExtra)
library(lattice)
library(lsr)
library(RColorBrewer)
library(tidyr)
library(vcd)

```

## Data Preparation

```{r preparation_BNC}
annotation_file <- read.csv(here("BNCspoken_prog_conc_anno.csv"), sep = "\t", header = TRUE, na.strings="")
glimpse(annotation_file)

# Select only annotated concordance lines
annotated <- annotation_file[71:3380,] # So far, I've only annotated to line 4006. This current selection ensures that we have the same number of progressives as in the Textbook Conversation subcorpus (namely 2423 progressives)

# Extract only progressives
BNCprog <- annotated[annotated$Tense != "NOT progressive" & annotated$Tense != "GOING TO" & annotated$Tense != "catenative", ]

BNCprog$Tense <- droplevels(BNCprog)$Tense # Drop unused levels
BNCprog$Time.reference <- droplevels(BNCprog)$Time.reference # Drop unused levels
BNCprog$Tense <- factor(BNCprog$Tense, levels = c("past", "perfect", "present", "modal")) # Re-order levels
BNCprog$Time.reference <- factor(BNCprog$Time.reference, levels = c("past", "past/present", "present", "future", "general", "hypothetical", "unclear"))

# Number of progressives extracted
nrow(BNCprog)
glimpse(BNCprog)


```

```{r preparation_TxB}
# Before saving annotation file as .csv, make sure to have deleted all trailing inverted commas as they serve as string deliminators. This can easily be done with a regex search and replace in calc $" --> nothing.
annotation_file_TxB <- read.csv(here("TxB_spoken_prog_conc_anno.csv"), sep = "\t", header = TRUE, na.strings="")
glimpse(annotation_file_TxB)

# Select only annotated concordance lines
annotated_TxB <- annotation_file_TxB

# Extract only progressives
TxBprog <- annotated_TxB[annotated_TxB$Tense != "NOT progressive" & annotated_TxB$Tense != "GOING TO" & annotated_TxB$Tense != "catenative" , ]

TxBprog$Tense <- droplevels(TxBprog)$Tense # Drop unused levels
TxBprog$Time.reference <- droplevels(TxBprog)$Time.reference # Drop unused levels

TxBprog$Tense <- factor(TxBprog$Tense, levels = c("past", "perfect", "present", "modal")) # Re-order levels
TxBprog$Time.reference <- factor(TxBprog$Time.reference, levels = c("past", "past/present", "present", "future", "general", "hypothetical", "unclear"))

# Number of progressives extracted
nrow(TxBprog)
# Ratio of false positives in progressive query
1 - nrow(TxBprog)/nrow(annotation_file_TxB)

```

```{r preparation_combine}

prog <- bind_rows("Spoken BNC2014 sample" = BNCprog, "Textbook Conversation" = TxBprog, .id = "Corpus")
glimpse(prog)
prog <- prog %>% dplyr::rename(Contraction = X., Question = X..1, Negation = NOT) # Rename variables
glimpse(prog)

# Replace NA with meaningful levels
summary(as.factor(prog$Negation))
prog$Negation <- tidyr::replace_na(as.character(prog$Negation), "positive") # positive
summary(as.factor(prog$Contraction))
prog$Contraction <- tidyr::replace_na(as.character(prog$Contraction), "full") # full
summary(as.factor(prog$Question))
prog$Question <- tidyr::replace_na(as.character(prog$Question), "statement") # statement

summary(as.factor(prog$Level))
prog$Level <- tidyr::replace_na(as.character(prog$Level), "Spoken BNC2014 sample")
prog$Textbook <- tidyr::replace_na(as.character(prog$Textbook), "Spoken BNC2014 sample")


# Make the following variables into factors
factor_cols <- c("Corpus", "Negation", "Contraction", "Question", "Repeated", "Lemma", "Voice", "Level")
prog[factor_cols] <- lapply(prog[factor_cols], as.factor)

# Drop unused levels
prog$Corpus <- droplevels(prog)$Corpus

levels(prog$Negation)[levels(prog$Negation)=="n"] <- "negated"
levels(prog$Contraction)[levels(prog$Contraction)=="c"] <- "contract."
levels(prog$Tense)[levels(prog$Tense)=="modal/infinitive"] <- "modal/inf."

summary(prog$Corpus) # Number of annotated progressive concordances in each subcorpus

#saveRDS(prog, file = here("prog.rds") last saved on 14/02/20
#saveRDS(prog, file = here("prog_collanalysis.rds")

```

```{r quick prog data access}

prog <- readRDS(file = here("prog.rds"))

```

# Proportion of progressives

```{r prop_prog}

# Textbook conversation #
TS_words <- 420130 # Much better than tokens due to lack of punctuation in Spoken BNC
TS_verbs <- 80106 # These are all verb forms (e.g. have been drinking = 3 verb forms). Calculated using SE query
TS_verb_phrases <- 64292 # As calculated with Extract_verb_lemmas_spacy
TS_prog <- nrow(TxBprog) #2423 # After manual annotation, thus excluding GOING TO + inf.
TS_ProgVerbPhrase_ratio <- TS_prog/(TS_verb_phrases - TS_prog - (nrow(filter(annotated_TxB, Tense=="GOING TO"))))*10000; TS_ProgVerbPhrase_ratio # 395.4756 per 10,000 finite verb phrases

# Spoken BNC
BNC_words <- 10581951
BNC_verbs <- 2487515
BNC_progCQL <- 126395
# Ratio of genuine progressives in progressive query for BNC
BNCratio_prog_CQL <- nrow(BNCprog)/nrow(annotated) # (using full BNC dataset as for coll.analysis)
BNCtotal_prog <- BNCratio_prog_CQL*BNC_progCQL # Approximate total number of progressives across BNC
BNC_verb_phrases <- 1751040
BNC_ProgVerbPhrase_ratio <- BNCtotal_prog/(BNC_verb_phrases - BNCtotal_prog - (nrow(filter(annotation_file, Tense=="GOING TO"))))*10000; BNC_ProgVerbPhrase_ratio # 566.0897 per 10,000 finite verb phrases (using full BNC dataset as for coll.analysis)

# Comparing relative frequencies #

TS <- as.vector(c(TS_words, TS_verbs, TS_verb_phrases, TS_prog))
BNC <- as.vector(c(BNC_words, BNC_verbs, BNC_verb_phrases, BNCtotal_prog))
BNC_sample <- round(BNC*(TS_words/BNC_words),0) # This has to be based on the total number of words because there are no sentence delimiters in the Spokem BNC 2014.
prog_ratio <- rbind(TS, BNC_sample)
rownames(prog_ratio) <- c("Textbook Conversation", "Spoken BNC2014 sample")
colnames(prog_ratio) <- c("Words", "Verb forms", "Finite verb phrases", "Progressives")
prog_ratio <- as.data.frame(prog_ratio); prog_ratio
prog_ratio$Non_prog <- prog_ratio$`Finite verb phrases` - prog_ratio$Progressives
prog_ratio$FProgRatio <- round(c(TS_ProgVerbPhrase_ratio, BNC_ProgVerbPhrase_ratio), 2)
colnames(prog_ratio) <- c("Words", "Verb forms", "Finite verb phrases", "Progressives", "Non-progressives", "F-coefficients of progressives")
prog_ratio
#clipr::write_clip(prog_ratio) # Copy into manuscript
#saveRDS(prog_ratio, file = here("prog_ratio_spoken.rds")

# Significance testing and approximation of effect sizes #
# http://rcompanion.org/handbook/H_03.html 

prog_ratio1 <- prog_ratio[,5:4]; prog_ratio1 # This is based on the word-based sample size of the Spoken BNC2014
chisq.test(prog_ratio1) # Highly significant
vcd::assocstats(t(prog_ratio1)) # Small effect size
fisher.test(prog_ratio1) # Is this helpful?

verb_ratio <- prog_ratio[,1:2]; verb_ratio
chisq.test(verb_ratio) # Highly significant
vcd::assocstats(t(verb_ratio)) # Very small effect size

# https://www.tutorialgateway.org/mosaic-plot-in-r/

#tiff(here("mosaic_prog_prop2.tiff"), height = 20, width = 30, units="cm", compression = "lzw", res = 300)

mosaicplot(prog_ratio1, shade = TRUE, main = "Progressives in Textbook Conversation and the Spoken BNC 2014", cex.axis = 1.5) # Unlike vcd::mosaic, Base R mosaicplot works with dataframes
#dev.off() 

```

# Tense, aspect and modality

```{r tense-forms}

# Table summaries
library(gtsummary)
# Great slides on using this package: http://www.emilyzabor.com/cleveland-r-gtsummary
# Also for adding custom functions: http://www.danieldsjoberg.com/gtsummary/reference/tests.html

prog %>% 
  select(Corpus, Tense) %>% 
  tbl_summary(by = Corpus) %>% 
  #add_p() %>% 
  modify_header(label = "**Form**") #%>% 
  #as_flex_table() %>%
  #flextable::save_as_docx(path = here("SpokenTensesTable.docx")

prog %>% 
  select(Tense, Level) %>% 
  tbl_summary(by = Level) %>% 
  #add_p() %>% 
  modify_header(label = "**Form**") #%>% 
  #as_flex_table() %>%
  #flextable::save_as_docx(path = here("SpokenTensesTable_byLevel.docx")

# Graphs with absolute values.
par(mfrow=c(1,2))
plot(BNCprog$Tense, main = "Progressives in the Spoken BNC2014 sample", ylab = "Number of occurrences", ylim = c(0,2000))
plot(TxBprog$Tense, main = "Progressives in the Spoken BNC2014 sample", ylab = "Number of occurrences", ylim = c(0,2000))
par(mfrow=c(1,1))

#png(filename = here("prog_tenses.png", width = 1800, height = 1400, res = 300)

# Distributions across corpora
round(prop.table(table(prog$Tense[prog$Corpus=="Textbook Conversation"])), 4)*100
round(prop.table(table(prog$Tense[prog$Corpus=="Spoken BNC2014 sample"])), 4)*100

# Mosaic plot of tense distribution across corpora
# Select the colours that will be used
library(RColorBrewer)
# All palette available from RColorBrewer
display.brewer.all(colorblindFriendly = T)
# We will select the first 4 colors in the Set1 palette
cols<-brewer.pal(n=4,name="OrRd")

#tiff(here("mosaic_tenseforms.tiff"), height = 30, width = 35, units="cm", compression = "lzw", res = 300)
par(cex = 2, cex.main = 0.85, font = 2)

#svg(here("mosaic_tenseforms.svg"), height = 10, width = 12, pointsize = 12)
par(cex = 2, cex.main = 0.85, font = 2)

plot(prog$Tense ~ prog$Corpus, col = cols, xlab = "", ylab = "Progressive forms", main = "Distribution of progressive forms in Textbook Conversation and the Spoken BNC2014")
#dev.off()

# Differentiated look at textbook levels #
#tiff(here("mosaic_tenseforms_levels.tiff"), height = 32, width = 40, units="cm", compression = "lzw", res = 300)
par(cex = 2)
plot(prog$Tense ~ relevel(prog$Level, "Spoken BNC2014 sample"), col = cols, xlab = "                                                   Textbook Levels", ylab = "")
#coord2 <- locator(4)
coord2 <- readRDS(file = here("coord2.rds"))
text(coord2, c("past", "perfect", "present", "modal"), font = 1)
#dev.off()

tenses <- table(prog$Tense, prog$Corpus)
tenses
tensesp <- t((prop.table(tenses, margin = 2)*100)) # Work out percentages and transpose table
tensesp <-round(tensesp, 2)

barplot(tensesp, beside = TRUE, ylab = "% of occurrences", xlab = "Forms of the progressive", ylim = c(0,100), col = c("darkblue", "darkred"))
legend("topright", fill = c("darkblue", "darkred"), c("BNC Spoken 2014", "Textbook Conversation"))
# Legend: "Forms of the progressives in the Spoken BNC 2014 and Textbook Conversation"

vcd::assocstats(tenses) # Measuring the effect size. Cramer's V is what we're interested in (Levshina p. 217)

#http://groups.chass.utoronto.ca/pol242/Labs/LM-3A/LM-3A_content.htm

chisq.test(tenses)
lsr::associationTest( ~ Tense + Corpus, data = prog) # Easier to interpret results

# Association plot of tense distribution in corpora
#tiff(here("tenseforms_assoc.tiff"), height = 12, width = 17, units="cm", compression = "lzw", res = 300)
vcd::assoc(tenses, shade = TRUE, varnames = FALSE) # This works
#dev.off()

chisq.test(tenses)$stdres

# Association plot of tense distributions across levels

levels(prog$Level) <- c("A", "B", "C", "D", "E", "Spoken BNC2014")
tenselevel <- table(prog$Tense, prog$Level) 

#tiff(here("tenseforms_levels_assoc.tiff"), height = 13, width = 20, units="cm", compression = "lzw", res = 300)
vcd::assoc(tenselevel, shade = TRUE, varnames = FALSE) # This works
#dev.off()

lsr::associationTest( ~ Tense + Level, data = prog) # Easier to interpret results

```

# Time reference

```{r time-ref}
# Distributions across corpora
round(prop.table(table(prog$Time.reference[prog$Corpus=="Textbook Conversation"])), 4)*100
round(prop.table(table(prog$Time.reference[prog$Corpus=="Spoken BNC2014 sample"])), 4)*100

timeref <- table(prog$Corpus, prog$Time.reference); timeref # Raw figures
timerefprop <- round(t((prop.table(timeref, margin = 2)*100)),2); timerefprop # As %

chisq.test(timeref)
vcd::assocstats(timeref) # Cramer's V = 0.263

# "Fairer" distribution, excluding "unclear's"
prog1 <- prog[prog$Time.reference!="unclear",]
prog1$Time.reference <- droplevels(prog1$Time.reference)
lsr::associationTest( ~ Corpus + Time.reference, data = prog1) # Cramer's V = 0.091

timeref <- table(prog1$Corpus, prog1$Time.reference); timeref # Raw figures

# Significance test for only textbook levels C-E against Textbook Conversation

prog2 <- prog[prog$Level!="A" & prog$Level!="B",]
prog2$Level <- droplevels(prog2)$Level # Drop levels A and B
prog2 <- prog2[prog2$Time.reference!="unclear",] # Get rid of unclear's
prog2$Time.reference <- droplevels(prog2)$Time.reference # Drop unclear level
timeref2 <- table(prog2$Corpus, prog2$Time.reference); timeref2 # Raw figures

chisq.test(timeref2) # p < 0.001
vcd::assocstats(timeref2) # Cramer's V = 0.153

#tiff(here("time-ref_assoc_levelsC-E.tiff"), height = 18, width = 36, units="cm", compression = "lzw", res = 300)
vcd::assoc(timeref2, shade = T, varnames = F, shading_max(), gp_labels=(grid::gpar(fontsize=14))) # Association plot only for textbook levels C-E, no unclear's
#dev.off()

# Association plot of time ref distribution in corpora

#tiff(here("time-ref_assoc.tiff"), height = 18, width = 36, units="cm", compression = "lzw", res = 300)
vcd::assoc(timeref, shade = TRUE, varnames = FALSE, gp_labels=(grid::gpar(fontsize=18)))
#dev.off()

# Barplot with number of occurrences per corpus
timeref <- table(prog$Corpus, prog$Time.reference); timeref # Raw figures
timeref <- as.data.frame(timeref)
names(timeref)[1] <- "Corpus"
names(timeref)[2] <- "Timeref"
names(timeref)[3] <- "Freq"
timeref

# Version 2 in a more publishable style
#tiff(here("timeref_barplot.tiff"), height = 15, width = 25, units="cm", compression = "lzw", res = 300)

cols<-RColorBrewer::brewer.pal(n=7,name="OrRd")

ggplot(timeref, aes(Timeref, Freq)) +
  geom_bar(aes(fill = Corpus), stat="identity", position="dodge", width=.5) +
  theme_bw(base_size = 18) +
  scale_fill_manual(values = cols[c(3,7)]) +
  ylab("Number of progressives") +
  xlab("Time reference") +
  scale_y_continuous(breaks = c(0, 250, 500, 750, 1000, 1250)) +
  theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_text(face = "plain", color = "grey20", size = 15), axis.text.y = element_text(face = "plain", color = "grey20", size = 15), legend.title=element_blank(), legend.position=c(.8, .8), legend.background = element_rect(colour="grey", linetype="solid"))
  #ggtitle("")

dev.off()

# Mosaic plot of time reference distribution across corpora

# We will select the first 4 colors in the Set1 palette
cols<-brewer.pal(n=7,name="OrRd")

#tiff(here("mosaic_timeref.tiff"), height = 40, width = 40, units="cm", compression = "lzw", res = 300)
par(cex = 2, cex.main = 0.85, font = 2)

#svg(here("mosaic_timeref.svg"), height = 10, width = 12, pointsize = 12)
par(cex = 2, cex.main = 0.85, font = 2)

plot(prog$Time.reference ~ prog$Corpus, col = cols, xlab = "", ylab = "Time reference of progressives", main = "Time reference of progressives in Textbook Conversation and the Spoken BNC2014 sample") # Good but bar plot is more informative
#dev.off()

# Differentiated look at textbook levels #
#tiff(here("mosaic_timeref_levels.tiff"), height = 30, width = 25, units="cm", compression = "lzw", res = 300)
par(cex = 1.5)
plot(prog$Time.reference ~ prog$Level, col = cols, xlab= "", ylab = "Time reference of the progressives")
mtext("                              Textbook levels", side = 1, adj = 0, line = 2.2, cex = 1.5) # To write text in the margin of plots
#dev.off()

# Differentiated look at textbook levels 2 #
prog1 <- prog[prog$Time.reference!="unclear",] # Get rid of unclear's
prog1$Time.reference <- droplevels(prog1)$Time.reference # Drop unclear level
cols<-brewer.pal(n=6,name="OrRd") # Pick just 6 colours now

#tiff(here("mosaic_timeref_levels.tiff"), height = 32, width = 30, units="cm", compression = "lzw", res = 300)
par(cex = 1.5)
plot(prog1$Time.reference ~ relevel(prog1$Level, "Spoken BNC2014"), col = cols, xlab = "                                                                    Textbook Levels", ylab = "", main = "Time reference of the progressive")
#coord <- locator(6)
coord <- readRDS(file = here("coord.rds"))
text(coord, c("past", "past/present", "present", "future", "general", "hypothetical"), font = 1)
#dev.off() # !! Crop graph to the left in Word afterwards

```

# Relationship between tense and time reference

```{r tense-time-ref}
# Plot relationship between tense and time reference

## Mosaic plot for comparison between two corpora ##

BNCprog$Tense <- relevel(BNCprog$Tense, "modal") # Makes it easier to read the mosaic plot if this tense form is first because it covers all time reference
TxBprog$Tense <- relevel(TxBprog$Tense, "modal")

par(mfrow=c(1,2))
plot(BNCprog$Time.reference ~ BNCprog$Tense)
plot(TxBprog$Time.reference ~ TxBprog$Tense) 
par(mfrow=c(1,1))

prog1 <- prog[prog$Time.reference != "unclear", ]
prog1$Time.reference <- droplevels(prog1)$Time.reference # Drop unused level "unclear"
prog1$Form <- prog1$Tense

cols<-brewer.pal(n=4,name="OrRd")
#tiff(here("mosaic_tenses_timeref.tiff"), height = 15, width = 28, units="cm", compression = "lzw", res = 300)
vcd::mosaic(Form ~ Corpus + Time.reference, prog1, gp = grid::gpar(fill = cols), rot_labels = c(0, 45, 0, 90), cex=2.5, zero_size = 0, labeling_args = list(rep = TRUE), xlab = "Time reference") # Best so far
dev.off()

vcd::assoc(Time.reference ~ Corpus + Tense, prog1, shade = TRUE) # Super hard to interpret!

```

```{r table-summaries-all-forms}

prog %>% 
  select(Corpus, Tense, Contraction, Question, Voice) %>% 
  tbl_summary(by = Corpus) %>% 
  add_p() %>% 
  modify_header(label = "**Form**") #%>% 
  # as_flex_table() %>%
  # flextable::save_as_docx(path = here("SpokenTensesTable.docx"))
  
prog %>% 
  select(Textbook, Contraction) %>% 
  tbl_summary(by = Textbook) %>% 
  modify_header(label = "**Form**") %>% 
  modify_footnote(update = everything() ~ NA)

prog %>% 
  select(Textbook, Negation) %>% 
  tbl_summary(by = Textbook) %>% 
  modify_header(label = "**Form**") %>% 
  modify_footnote(update = everything() ~ NA)

prog %>% 
  select(Corpus, Negation) %>% 
  tbl_summary(by = Corpus) %>% 
  modify_header(label = "**Form**")

prog %>% 
  select(Corpus, Question) %>% 
  tbl_summary(by = Corpus) %>% 
  modify_header(label = "**Form**")

prog %>% 
  select(Corpus, Voice) %>% 
  tbl_summary(by = Corpus) %>% 
  modify_header(label = "**Form**")

prog %>% 
  select(Corpus, Time.reference) %>% 
  tbl_summary(by = Corpus) %>% 
  modify_header(label = "**Time reference**") 

prog %>% 
  select(Corpus, Repeated) %>% 
  tbl_summary(by = Corpus) %>% 
  modify_header(label = "**Function**") 

prog1 <- prog
prog1$Extra.function <- factor(prog1$Extra.function, exclude = NULL)
levels(prog1$Extra.function)[is.na(levels(prog1$Extra.function))] <- "none"
prog1 %>% 
  select(Corpus, Extra.function) %>% 
  tbl_summary(by = Corpus, missing = "ifany") %>% 
  modify_header(label = "**Function**") 

prog %>% 
  select(Corpus, Continuous) %>% 
  tbl_summary(by = Corpus) %>% 
  modify_header(label = "**Function**") 

```

# Questions

H0: The frequencies of the levels of the dependent variable QUESTION do not vary as a function of the levels of the independent variable CORPUS; X2 = 0

H1: The frequencies of the levels of the dependent variable QUESTION vary as a function of the levels of the independent variable CORPUS; X2 \> 0

A chi-square test assumes independent observations which we don't really have here (lots from same textbook, possibly several from individual speaker in the BNC).

```{r questions}

round(prop.table(table(prog$Corpus, prog$Question), margin = 1), 4)*100

questions <- table(prog$Corpus, prog$Question)
questions <- questions[,1:2] # Get rid of 5 unclear's in the BNC data
test.questions <- chisq.test(questions); test.questions # Significant correlation at p < 0.001

assocplot(t(questions))
vcd::assoc(questions, shade = TRUE, col_vars = c("question", "statement"))
vcd::assocstats(questions) # Odd ratio of 32.571. Phi-coefficient is only 0.092 = very weak effect!

prog1 <- prog[prog$Question!="unclear",]
prog1$Question <- droplevels(prog1$Question)
lsr::associationTest( ~ Question + Corpus, data = prog1) # Cramer's V = 0.091

```

# Negation

H0: The frequencies of the levels of the dependent variable NEGATION do not vary as a function of the levels of the independent variable CORPUS; X2 = 0

H1: The frequencies of the levels of the dependent variable NEGATION vary as a function of the levels of the independent variable CORPUS; X2 \> 0

A chi-square test assumes independent observations which we don't really have here (lots from same textbook, possibly several from indidivudal speaker in the BNC).

```{r negation}

round(prop.table(table(prog$Corpus, prog$Negation), margin = 1), 4)*100

prog$Negation <- relevel(prog$Negation, "positive")

#tiff(here("prog_negation.tiff"), height = 17, width = 20, units="cm", compression = "lzw", res = 300)
par(cex = 1.4, cex.main = 0.85)
plot(prog$Negation ~ prog$Corpus, ylab = "", xlab = "", main = "", col = cols[3:4])
dev.off()

# No interesting correlation to be seen
#tiff(here("prog_negation_question.tiff", height = 13, width = 20, units="cm", compression = "lzw", res = 300)
par(cex = 1.4, cex.main = 0.85)
vcd::mosaic(Negation ~ Corpus + Question, prog, gp = grid::gpar(fill = cols[4:3]), rot_labels = c(0, 90, 0, 90), cex=2.5, zero_size = 0, labeling_args = list(rep = TRUE)) # Best so far
#dev.off()

negation <- table(prog$Corpus, prog$Negation)

test.negation <- chisq.test(negation); test.negation # Significant correlation at p < 0.05

assocplot(t(negation))
vcd::assocstats(negation) # Phi-coefficient is only 0.03 = very weak!

lsr::associationTest( ~ Negation + Corpus, data = prog) # Cramer's V = 0.03

```

# Contraction

```{r contraction}

round(prop.table(table(prog$Corpus, prog$Contraction), margin = 1), 4)*100
round(prop.table(table(prog$Textbook, prog$Contraction), 1),4)*100
table(prog$Textbook, prog$Contraction)

cols<-RColorBrewer::brewer.pal(n=4,name="OrRd")

#tiff(here("prog_contractions.tiff"), height = 25, width = 20, units="cm", compression = "lzw", res = 300)
par(cex = 1.4, cex.main = 0.85)
plot(prog$Contraction ~ prog$Corpus, ylab = "", xlab = "", main = "", col = cols[4:3])

plot(prog$Contraction ~ prog$Corpus, ylab = "", xlab = "", main = "", col = cols[4:3])

#dev.off()

#tiff(here("prog_contractions_tenses.tiff"), height = 13, width = 20, units="cm", compression = "lzw", res = 300)
par(cex = 1.4, cex.main = 0.85)
vcd::mosaic(Contraction ~ Corpus + Tense, prog, gp = grid::gpar(fill = cols[4:3]), rot_labels = c(0, 90, 0, 90), cex=2.5, zero_size = 0, labeling_args = list(rep = TRUE)) # Best so far
#dev.off()

round(prop.table(table(prog$Contraction, prog$Textbook), 2),4)*100

#tiff(here("prog_contractions_series.tiff"), height = 13, width = 20, units="cm", compression = "lzw", res = 300)
par(cex = 1.4, cex.main = 0.85)
vcd::mosaic(Contraction ~ Textbook, prog, gp = grid::gpar(fill = cols[4:3]), rot_labels = c(0, 90, 0, 90), cex=2.5, zero_size = 0, labeling_args = list(rep = TRUE)) # Best so far
#dev.off()

contractions <- table(prog$Corpus, prog$Contraction)
test.contractions <- chisq.test(contractions)
test.contractions # Significant correlation

assocplot(t(contractions))
vcd::assocstats(contractions) # Phi-coefficient is only 0.066 = very weak!

```

# Voice

```{r voice}

round(prop.table(table(prog$Corpus, prog$Voice), margin = 1), 4)*100

voice <- table(prog$Corpus, prog$Voice); voice
voice1 <- voice[,c(1:2)]; voice1 # Get rid of unclear as otherwise chi-squared approximation may be incorrect 
chisq.test(voice1) # p-value = 0.1863 = Not a significant correlation
vcd::assocstats(voice) # Phi-coefficient is only 0.026 = no effect!

prog[prog$Voice=="P",c(1,6,11)] # For qualitative analysis
table(prog[prog$Voice=="P",c(1,11)])

prog1 <- prog[prog$Voice!="unclear",]
prog1$Voice <- droplevels(prog1$Voice)
lsr::associationTest( ~ Voice + Corpus, data = prog1) # Cramer's V = 0.02

```

# Repeatedness

```{r repeatedness}

repeatedness <- table(prog$Corpus, prog$Repeated); repeatedness
round(prop.table(repeatedness, 1),4)*100
chisq.test(repeatedness) # Significant correlation at p<0.001
vcd::assocstats(repeatedness)

#tiff(here("Prog_Repeatedness.tiff"), height = 18, width = 20, units="cm", compression = "lzw", res = 300)
par(cex = 1.5, cex.main = 1)
par(cex = 1, cex.lab = 1.5, cex.axis = 1, cex.main = 1)

plot(repeatedness, shade = TRUE, main = "Progressive function: Repeatedness")
#dev.off()

repeatedness1 <- repeatedness[,c(1,3)] # Get rid of unclear's - if so desired... (I'm not sure this is a good idea because they make up some 9% of BNC data)
chisq.test(repeatedness1)
vcd::assocstats(repeatedness1) # Phi-coefficient is only 0.082 = very weak!

## Which lemmas occur most frequently in either corpora with the function repeatedness? 
repeated_lemmas <- prog[prog$Repeated=="yes", c(1,8)] 
repeated_lemmas_TxB <- prog[prog$Repeated=="yes" & prog$Corpus=="Textbook Conversation", 8] 
table_repeated_lemmas_TxB <- sort(table(repeated_lemmas_TxB), decreasing = T)
head(table_repeated_lemmas_TxB, 10)

repeated_lemmas_BNC <- prog[prog$Repeated=="yes" & prog$Corpus=="Spoken BNC2014 sample", 8] 
table_repeated_lemmas_BNC <- sort(table(repeated_lemmas_BNC), decreasing = T)
head(table_repeated_lemmas_BNC, 10)

## Which lemmas occur proportionally most frequently with function repeatedness?
BNCrepeated <- BNCprog[, c(7, 15)]
BNCrepeated <- table(Case = BNCrepeated$Lemma, BNCrepeated$Repeated)
BNCrepeated <- addmargins(BNCrepeated) # adds rows total to table!
BNCrepeated <- BNCrepeated[order(BNCrepeated[,4], decreasing = T),]; head(BNCrepeated, 11)
BNCrepeated <- BNCrepeated[1:30,]
BNCrepeatedp <- round(prop.table(BNCrepeated, 1), 4)*200; BNCrepeatedp
BNCrepeatedp <- BNCrepeatedp[order(BNCrepeatedp[,3], decreasing = T),]
BNCrepeatedp <- BNCrepeatedp[1:12,1:3]; BNCrepeatedp

fig_BNC_repeated = ggplot(as.data.frame(BNCrepeatedp), aes(x=Var2, y=Freq)) +
  geom_col() +
  facet_wrap(~Case) +
  labs(x='Repeatedness function', y='Proportion of use') +
  theme(axis.text.x=element_text(angle=45, hjust=1))
print(fig_BNC_repeated) # Takes a few seconds

TCrepeated <- TxBprog[, c(7, 15)]
TCrepeated <- addmargins(table(Case = TCrepeated$Lemma, TCrepeated$Repeated)); TCrepeated
TCrepeated <- TCrepeated[order(TCrepeated[,4], decreasing = T),]; head(TCrepeated, 11)
TCrepeated <- TCrepeated[1:30,]
TCrepeatedp <- round(prop.table(TCrepeated, 1), 4)*200; TCrepeatedp
TCrepeatedp <- TCrepeatedp[order(TCrepeatedp[,3], decreasing = T),]
TCrepeatedp <- TCrepeatedp[1:9,1:3]; TCrepeatedp

fig_TC_repeated = ggplot(as.data.frame(TCrepeatedp), aes(x=Var2, y=Freq)) +
  geom_col() +
  facet_wrap(~Case) +
  labs(x='Repeatedness function', y='Proportion of use') +
  theme(axis.text.x=element_text(angle=45, hjust=1))
print(fig_TC_repeated) # Takes a few seconds

##

rep <- prog[, c(1, 8, 16)]; head(rep)
rep <- table(Case = rep$Lemma, rep$Repeated)
BNCrepeated <- addmargins(BNCrepeated) # adds rows total to table!
BNCrepeated <- BNCrepeated[order(BNCrepeated[,4], decreasing = T),]; head(BNCrepeated, 11)
BNCrepeated <- BNCrepeated[1:30,]
BNCrepeatedp <- round(prop.table(BNCrepeated, 1), 4)*200; BNCrepeatedp
BNCrepeatedp <- BNCrepeatedp[order(BNCrepeatedp[,3], decreasing = T),]
BNCrepeatedp <- BNCrepeatedp[1:12,1:3]; BNCrepeatedp

```

# Continuousness

```{r continousness}

continuous <- table(prog$Corpus, prog$Continuous); continuous
prop.table(continuous, 1)*100

#tiff(here("Prog_Continuous.tiff"), height = 18, width = 20, units="cm", compression = "lzw", res = 300)
par(cex = 1.6, cex.main = 0.6, cex.lab = 1, cex.axis = 1.5, cex.sub = 0.5)
plot(continuous, shade = TRUE, main = "Progressive function: Continuousness")
dev.off()

continuous1 <- continuous[,c(1,3)] # Get rid of unclear's
chisq.test(continuous1) # Significant correlation at p<0.001
vcd::assocstats(continuous1) # Cramer's V is 0.147

## Which lemmas occur most frequently in either corpora with the function non-continuousness? ##
noncont_lemmas <- prog[prog$Continuous=="no", c(1,8)] 
noncont_lemmas_TxB <- prog[prog$Continuous=="no" & prog$Corpus=="Textbook Conversation", 8] 
table_noncont_lemmas_TxB <- sort(table(noncont_lemmas_TxB), decreasing = T)
head(table_noncont_lemmas_TxB, 20)

noncont_lemmas_BNC <- prog[prog$Continuous=="no" & prog$Corpus=="Spoken BNC2014 sample", 8] 
table_noncont_lemmas_BNC <- sort(table(noncont_lemmas_BNC), decreasing = T)
head(table_noncont_lemmas_BNC, 20)

## Percentage of specific lemmas used with continuousness function

BNCPC <- BNCprog[BNCprog$Continuous!="unclear",]
prop.table(table(BNCPC$Lemma=="tell", BNCPC$Continuous), 1)*100

TCPC <- TxBprog[TxBprog$Continuous!="unclear",]
prop.table(table(TCPC$Lemma=="tell", TCPC$Continuous), 1)*100

```

# Additional functions

```{r framing}

functions1 <- table(prog$Corpus, prog$Extra.function); functions1
round((functions1/2423*100),2) # Percentage of concordances in each corpus

framing <- table(prog$Corpus, prog$Extra.function=="framing"); framing
chisq.test(framing) # Significant at p>0.001
vcd::assocstats(framing) # phi-coefficient = 0.242

emphasis <- table(prog$Corpus, prog$Extra.function=="emphasis/shock"); emphasis
chisq.test(emphasis) # Not significant
vcd::assocstats(emphasis)

change <- table(prog$Corpus,prog$Extra.function=="gradual change"); change
chisq.test(change) # Significant at p < 0.005
vcd::assocstats(change) # Phi-Coefficient = 0.171

politeness <- table(prog$Corpus, prog$Extra.function=="politeness/softening"); politeness
chisq.test(politeness) # 
vcd::assocstats(politeness) # 

```

